---
title: "gbmRegressionResults"
author: "Ian Douglas"
date: "5/23/2020"
output: html_document
---
```{r}
library(gbm)
library(tidyverse)
library(MLmetrics)
library(haven)
source("getBestTreeAndRMSE.R")
source("getBestTreeAndR2.R")
source("getBestTreePredictions.R")
```

# Read in the final results from the server
```{r}
gbmReg_res <- readRDS("output/gbmRegrLongData_2020-06-02.rds")
```
# Read in the data and process identically to how it was for the analysis
```{r}
strDat <- readRDS("../../data/master/noCC-master-StrDataLong_GAM-Adjusted_2020-05-20.rds") %>%
  mutate_at(vars(GROUP), ~as.numeric(. == "PI")) %>%
  dplyr::select(-GENDER_FEMALE, -brain_age_yrs, -EstimatedTotalIntraCranialVol)
##### Note-retained colums: subjWeight, IDENT_SUBID
# Read in the cbcl (starting with the SB master df)
sb.master <- haven::read_sav("../../data/master/J.Data_Entry_Master_8_17_17_fixed_fixed_4.sav")
cbcl <- sb.master %>% # from the data frame in which the cbcl survey is the older kids
  select(IDENT_SUBID, contains("SUBAGE_s", ignore.case = F), contains("CBCL_")) %>%
  select(IDENT_SUBID, contains("SUBAGE_s"), contains("INTERN"), contains("EXTERN"), contains("TOTPROB")) %>%
  select(IDENT_SUBID, contains("SUBAGE_s"), contains("PROB_T"), contains("PROB_TOT")) %>%
  # reorder the columns
  select(1:4, matches("_T$"), matches("TOT$"), contains("FU1"), contains("FU2"), everything()) %>%
  # extend the ages down into long form & use them to create a WAVE column to match that in the brain data
  gather(., key = 'WAVE', value = 'brain_age_yrs', SUBAGE_session2:SUBAGE_session4) %>%
  # convert it to yrs
  mutate_at('brain_age_yrs', ~ ./12) %>%
  select(IDENT_SUBID, WAVE, brain_age_yrs, everything()) %>% # reorder the columns again
  # now change the subage_session2 to WAVE2, and so forth for session3 and 4; so, (subage_session2 --> WAVE1)
  mutate_at("WAVE", ~paste0("WAVE", as.numeric(sub("SUBAGE_session", "", .)) - 1)) %>%
  arrange(IDENT_SUBID, WAVE)
# Collapse the entries for each timepoint from the cbcl_1.5 and cbcl_4 for each category
INTRNLZNG_T <- INTRNLZNG_TOT <- EXTRNLZNG_T <- EXTRNLZNG_TOT <- TOTALPROB_T <- TOTALPROB_TOT <- NULL
# define quick helper func
get.dat <- function(pat1, pat2)
{
  na.omit(unlist(
    cbclDat[i, grepl(pat1, names(cbclDat)) & grepl(pat2, names(cbclDat))]
  ))
}

for (i in 1:nrow(cbcl)) {
  if (cbcl$WAVE[i] == "WAVE1") {
    cbclDat <- cbcl %>% select(ends_with("T")) # either "T" or "TOT" desired
  } else if (cbcl$WAVE[i] == "WAVE2") {
    cbclDat <- cbcl %>% select(ends_with("_FU1"))
  } else cbclDat <- cbcl %>% select(ends_with("_FU2"))
  
  res <- get.dat("INTERNPROB", "_T$|_T_")
  INTRNLZNG_T[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("INTERNPROB", "_TOT$|_TOT_")
  INTRNLZNG_TOT[i] <- ifelse(length(res) == 1, res, NA) # not used below
  res <- get.dat("EXTERNPROB", "_T$|_T_")
  EXTRNLZNG_T[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("EXTERNPROB", "_TOT$|_TOT_")
  EXTRNLZNG_TOT[i] <- ifelse(length(res) == 1, res, NA) # not used below
  res <- get.dat("TOTPROB", "_T$|_T_")
  TOTALPROB_T[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("TOTPROB", "_TOT$|_TOT_")
  TOTALPROB_TOT[i] <- ifelse(length(res) == 1, res, NA) # not used below
}
# Finally, extract the CBCL variable in long format
# Just taking the T scores
cbcl_long = data.frame(
  dplyr::select(.data = cbcl, IDENT_SUBID, WAVE),
  CBCL_INTERNALIZING_T = INTRNLZNG_T, # just the t-scores
  CBCL_EXTERNALIZING_T = EXTRNLZNG_T,  # just the t-scores
  CBCL_TOTALPROB_T = TOTALPROB_T,  # just the t-scores
  stringsAsFactors = F
)
# Merge the CBCL and SGMV data frame
cbclAndBrain <- left_join(x = strDat, y = cbcl_long,
                          by = c("IDENT_SUBID", "WAVE"),
                          all = T) %>%
  select(IDENT_SUBID:WAVE, subjWeight, everything()) %>% ungroup()

# Finally, seperate the data frame into a list of data frames, 1 for each cbcl scale
# Also rename it CBCL
data.list <- lapply(grep("^CBCL",names(cbclAndBrain),value = T), function(nm) {
  cbclAndBrain %>%
    # select from the first variable, up to everything except the CBCL variables
    # But do not "de-select" cbcl variables, because we need just the one in "nm"
    select(1:(grep("CBCL", names(.))[1]-1), nm) %>% # cbcl is the last col.
    rename("CBCL" = last_col()) %>%
    drop_na(last_col()) # drop_na with respect to the cbcl column
})
names(data.list) <- grep("^CBCL",names(cbclAndBrain),value = T)
```

# print the number of resampling iterations
```{r}
lapply(gbmReg_res, length)
```

# Summarize the results
```{r}
gbmResFrame.list <- lapply(gbmReg_res, function(x) {
  data.frame(
    cross.val.RMSE = sapply(x, function(xx) xx$GBM$BestModel_TestRMSE), # retains rownames
    n.trees = sapply(x, function(xx) xx$GBM$BestModel_BestNumTrees),
    map_dfr(.x = x, .f = function(xx) xx$GBM$BestParams)
  )}
)
```

# print the mean across the cross validation iterations
```{r}
lapply(1:length(gbmResFrame.list), function(i) {
  summarize(gbmResFrame.list[[i]], meanRMSE = mean(cross.val.RMSE), sdRMSE = sd(cross.val.RMSE))
})
```

# Plot the distribution of the RMSEs
```{r}
# lapply(gbmResFrame.list, function(df) {
#   ggplot(rownames_to_column(gbmResFrame.list, "resample") %>%
#            mutate(meanAUC = mean(cross.val.RMSE))) +
#     geom_histogram(aes(x = cross.val.RMSE), bins = 20, color = 'black', fill = 'lightgrey') +
#     stat +
#     theme_linedraw()
# })
lapply(names(gbmResFrame.list), function(nm) {
  rownames_to_column(gbmResFrame.list[[nm]], "resample") %>%
    mutate(meanRMSE = mean(cross.val.RMSE)) %>%
    ggplot(.) +
    geom_histogram(aes(x = cross.val.RMSE), bins = 20, color = 'black', fill = 'lightgrey') +
    geom_vline(aes(xintercept = meanRMSE[1]), linetype = 2) +
    theme_linedraw() +
    ggtitle(paste0("Distribution of Cross Validation RMSE Computed on the Heldout Test Set\n for 100 Gradient Boosted Machines Predicting ", nm)) + 
    theme(plot.title = element_text(size = 20, hjust = .5), 
          axis.title = element_text(size = 18), 
          axis.text = element_text(size = 15)) +
    xlab("RMSE")
})
```
# Get the permutation p-values of each cross-validation iteration
```{r}
pvals = map(names(gbmResFrame.list), function(x) { # for each outome
  sapply(rownames(gbmResFrame.list[[x]]), function(xx) { # for each resample
    y_hat = gbmReg_res[[x]][[xx]]$GBM$BestModelTestSetPreds # get the preds
    y_perms = lapply(1:1000, function(i) { # 1000 times, permute the test set TRUE outcome
      sample(gbmReg_res[[x]][[xx]]$data$true_y)
    })
    nullRMSE = sapply(y_perms, function(yperm) {RMSE(y_hat, yperm)})
    p = (sum(gbmReg_res[[x]][[xx]]$GBM$BestModel_TestRMSE > nullRMSE) + 1) / (length(nullRMSE) + 1)
    p
  })
})
gbmResFrame.list[[1]]$pvalue <- pvals[[1]]
gbmResFrame.list[[2]]$pvalue <- pvals[[2]]
gbmResFrame.list[[3]]$pvalue <- pvals[[3]]
# re-save
saveRDS(gbmResFrame.list, "results/regression/2020-06-02/gbmResults_FINAL-LIST_2020-06-02.rds")
lapply(gbmResFrame.list, head)
```
# Investigation of the hyperparameters
```{r}
lapply(names(gbmResFrame.list), function(nm) {
  rownames_to_column(gbmResFrame.list[[nm]], "resample") %>%
    mutate_at("resample", ~as.numeric(sub("R.+_", "", .))) %>%
    pivot_longer(-one_of("resample", "cross.val.RMSE", "pvalue", "n.trees"), names_to = "Hyperparameter",values_to= "Optimal.Value") %>%
    arrange(resample, Optimal.Value) %>%
    mutate(Significance = as.logical(pvalue < .05)) %>%
    ggplot(data = ., aes(x = factor(Optimal.Value), y = cross.val.RMSE)) +
    geom_jitter(aes(group = Hyperparameter, color = Significance), alpha = .9, width = .1) +
    stat_summary(fun.data = "mean_cl_boot", colour = "black") + 
    facet_wrap(~Hyperparameter, scales = 'free') +
    theme(plot.title = element_text(size = 20, hjust = .5),
          axis.text.x = element_text(size = 15, angle = 45, hjust = .85, vjust = 1),
          axis.text.y = element_text(size = 15)) +
    labs(title = "Prediction RMSE of the Optimal Models and the Final \nValues of their Tuning Parameters") +
    xlab("Optmal Hyperparameter Value") + ylab(paste0("Test RMSE in predicting ", nm)) + 
    scale_color_discrete() + 
    guides(color = guide_legend(reverse = T))
  ggsave(paste0("results/regression/2020-06-02/hyperparamFinalValues-",nm,".jpg"), units = "in",
         height = 6, width = 11, device = "jpg")
  
})
```

# Visualize the distribution of the p-values
```{r}
lapply(names(gbmResFrame.list), function(nm) {
  rownames_to_column(gbmResFrame.list[[nm]], "resample") %>%
    mutate(meanpval = mean(pvalue)) %>%
    ggplot(.) +
    geom_vline(xintercept = .05, linetype = 1, color = "blue") +
    geom_histogram(aes(x = pvalue), bins = 100, color = 'black', fill = 'lightgrey') +
    geom_vline(aes(xintercept = meanpval[1]), linetype = 2) +
    theme_linedraw() +
    ggtitle(paste0("Distribution of Permutation p-values Associated with the Test-set RMSE of each GBM model\n for the Predicted Outcome ", nm)) + 
    labs(caption = "Dotted line denotes the mean\nBlue line indicates alpha at .05") +
    theme(plot.title = element_text(size = 20, hjust = .5), 
          axis.title.y = element_text(size = 18), 
          axis.text = element_text(size = 15),
          axis.title.x = element_text(size = 18, face = "italic")) +
    xlab("p")
  ggsave(paste0("results/regression/2020-06-02/pValues-Predicting_",nm,"_EachModelIndividualPermTest.jpg"),
         units = "in", height = 6.6, width = 11, device = "jpg")
})

```

# Obtain the average participant's prediction for when they were out of sample
```{r}
testFrame.list = lapply(names(data.list), function(nm) {
  cbind(data.list[[nm]] %>% dplyr::select(IDENT_SUBID) %>% filter(!duplicated(IDENT_SUBID)),
        as.data.frame(lapply(rownames(gbmResFrame.list[[nm]]), function(r) {
          y_hatFrame = data.frame(IDENT_SUBID = gbmReg_res[[nm]][[r]]$data$test.subid,
                                  y_hat = gbmReg_res[[nm]][[r]]$GBM$BestModelTestSetPreds,
                                  stringsAsFactors = F) %>%
            group_by(IDENT_SUBID) %>%
            # average for participants who were in the test set more than once
            summarize(y_hat = mean(y_hat)) %>% ungroup()
          merged = merge(data.list[[nm]] %>% dplyr::select(IDENT_SUBID) %>%
                           filter(!duplicated(IDENT_SUBID)),
                         y_hatFrame, by = "IDENT_SUBID", all.x = TRUE, all.y = F)
          return(merged$y_hat)
        })))
})
names(testFrame.list) <- names(data.list)
#
averageTestSetPreds.list = lapply(testFrame.list, function(df) {
  data.frame(
    IDENT_SUBID = df$IDENT_SUBID,
    predicted = rowMeans(dplyr::select(df, -IDENT_SUBID), na.rm = T),
    predSE = unlist(apply(dplyr::select(df, -IDENT_SUBID), 1, function(x) {
      sd(x, na.rm = T)/sqrt(sum(!is.na(x)))
    })),
    n = rowSums(!is.na(dplyr::select(df, -IDENT_SUBID))),
    stringsAsFactors = F
  ) %>%
  select(IDENT_SUBID, predicted, predSE, n)
})
#
ResultsFrame.list = map(names(averageTestSetPreds.list), function(nm) {
  merge(averageTestSetPreds.list[[nm]],
        data.list[[nm]] %>% dplyr::select(IDENT_SUBID, CBCL) %>%
          filter(!duplicated(IDENT_SUBID)),
        by = "IDENT_SUBID") %>%
    mutate(error = CBCL - predicted)
})
names(ResultsFrame.list) <- names(averageTestSetPreds.list)
saveRDS(testFrame.list, 
          "results/regression/2020-06-02/sparseAllResponses-RawTestSetPredictionsBySubject.rds")
saveRDS(ResultsFrame.list, 
          "results/regression/2020-06-02/FINAL-aggregateCrossValPredsAndScore-LIST-allCBCLResponses.rds")
```
# Performance of the averaged models:
```{r}
avgRMSE.list = map(ResultsFrame.list, function(x) RMSE(y_pred = x$predicted, y_true = x$CBCL))
avgRMSE.list
```
# Permutation significance of the average prediction model
```{r}
nullRMSEs.list = lapply(ResultsFrame.list, function(x) {
  sapply(1:1000, function(i) RMSE(x$predicted, sample(x$CBCL)))
})
averageModPval.list = lapply(1:length(avgRMSE.list), function(i) {
  (sum(avgRMSE.list[[i]] > nullRMSEs.list[[i]]) + 1) / (length(nullRMSEs.list[[i]]) + 1)
})
averageModPval.list
```
```{r}
pseudoR2.list <- lapply(1:3, function(i) {
  1 - (avgRMSE.list[[i]]/sqrt(var(ResultsFrame.list[[i]]$CBCL)))
})
pseudoR2.list
```

```{r}
jpeg("results/regression/2020-06-02/permTestRes_InternalFINAL-MODEL_RMSE.jpg",
     units = "px", height = 650, width = 1300)
plot(density(x = nullRMSEs.list[[1]]), main = "Null Distribution of RMSE values\nfor the Model of Internalizing CBCL",
     xlim = c(11, 12.7), sub = "Dotted line indicates Observed RMSE")
abline(v = avgRMSE.list$CBCL_INTERNALIZING_T, lty = 2)
dev.off()
plot(density(x = nullRMSEs.list[[1]]), main = "Null Distribution of RMSE values\nfor the Model of Internalizing CBCL",
     xlim = c(11, 12.7), sub = "Dotted line indicates Observed RMSE")
abline(v = avgRMSE.list$CBCL_INTERNALIZING_T, lty = 2)
```
```{r}
jpeg("results/regression/2020-06-02/permTestRes_ExternalFINAL-MODEL_RMSE.jpg",
     units = "px", height = 650, width = 1300)
plot(density(x = nullRMSEs.list[[2]]), main = "Null Distribution of RMSE values\nfor the Model of Externalizing CBCL",
     xlim = c(12, 13), sub = "Dotted line indicates Observed RMSE")
abline(v = avgRMSE.list$CBCL_EXTERNALIZING_T, lty = 2)
dev.off()
plot(density(x = nullRMSEs.list[[2]]), main = "Null Distribution of RMSE values\nfor the Model of Externalizing CBCL",
     xlim = c(12, 13), sub = "Dotted line indicates Observed RMSE")
abline(v = avgRMSE.list$CBCL_EXTERNALIZING_T, lty = 2)
```
```{r}
jpeg("results/regression/2020-06-02/permTestRes_TotalFINAL-MODEL_RMSE.jpg",
     units = "px", height = 650, width = 1300)
plot(density(x = nullRMSEs.list[[3]]), main = "Null Distribution of RMSE values\nfor the Model of Total CBCL",
     xlim = c(12.9, 14.4), sub = "Dotted line indicates Observed RMSE")
abline(v = avgRMSE.list$CBCL_TOTALPROB_T, lty = 2)
dev.off()
plot(density(x = nullRMSEs.list[[3]]), main = "Null Distribution of RMSE values\nfor the Model of Total CBCL",
     xlim = c(12.9, 14.4), sub = "Dotted line indicates Observed RMSE")
abline(v = avgRMSE.list$CBCL_TOTALPROB_T, lty = 2)
```

# Variable Importances
Check to make sure all the variable importance computations output the variables in the same order.
```{r}
allTheSameOrder1 = FALSE
current = rownames(gbmReg_res$CBCL_INTERNALIZING_T$Resample_1$PVI)
for (i in 2:length(gbmReg_res$CBCL_INTERNALIZING_T)) {
  target = rownames(gbmReg_res$CBCL_INTERNALIZING_T[[i]]$PVI)
  if (all.equal(target, current)) {
    current <- target
  } else break
  if (i == length(gbmReg_res$CBCL_INTERNALIZING_T)) allTheSameOrder1 <- TRUE 
}
allTheSameOrder2 = FALSE
current = rownames(gbmReg_res$CBCL_EXTERNALIZING_T$Resample_1$PVI)
for (i in 2:length(gbmReg_res$CBCL_EXTERNALIZING_T)) {
  target = rownames(gbmReg_res$CBCL_EXTERNALIZING_T[[i]]$PVI)
  if (all.equal(target, current)) {
    current <- target
  } else break
  if (i == length(gbmReg_res$CBCL_EXTERNALIZING_T)) allTheSameOrder2 <- TRUE 
}
allTheSameOrder3 = FALSE
current = rownames(gbmReg_res$CBCL_TOTALPROB_T$Resample_1$PVI)
for (i in 2:length(gbmReg_res$CBCL_TOTALPROB_T)) {
  target = rownames(gbmReg_res$CBCL_TOTALPROB_T[[i]]$PVI)
  if (all.equal(target, current)) {
    current <- target
  } else break
  if (i == length(gbmReg_res$CBCL_TOTALPROB_T)) allTheSameOrder3 <- TRUE 
}
c(allTheSameOrder1, allTheSameOrder2, allTheSameOrder3)
```
Proceed if TRUE TRUE TRUE.

Define a function to bind up the variable importances by their columns
```{r}
matchAndBindCol = function(df1, df2)
{
  df2_arranged = df2[match(df1$variable,df2$variable) ,]
  if (all.equal(df1$variable, df2$variable)) {
    return(cbind(df1, df2))
  }
}
varImpFrame.list <- lapply(gbmReg_res, function(res_list) {
  Reduce(f = matchAndBindCol, x = lapply(res_list, function(x) {
  data.frame(variable = rownames(x$PVI), importance = rowMeans(x$PVI), row.names = 1)
  })) %>%
    rownames_to_column("variable") %>%
    mutate(meanImp = rowMeans(dplyr::select(., starts_with("imp")))) %>%
    arrange(meanImp) %>%
    mutate(variable = factor(variable, levels = variable)) %>%
    dplyr::select(starts_with("imp"), everything()) %>%
    setNames(., nm = c(paste0('importance', 1:100), names(.)[-1:-100])) %>%
    mutate(avgRank = rowMeans(
      mutate_all(.tbl = dplyr::select(., starts_with("imp")), ~dense_rank(desc(.))))) %>%
    mutate(
      avgAbsRank = rowMeans(mutate_all(.tbl = dplyr::select(., starts_with("imp")), ~dense_rank(desc(abs(.)))))
    ) %>%
    dplyr::select(variable, meanImp, avgRank, avgAbsRank, everything())
})
saveRDS(varImpFrame.list,"results/regression/2020-06-02/VARIMP-LIST_finalRegressionResults-2020-05-28.rds")
```

# Plot var imps
```{r}
lapply(names(varImpFrame.list), function(nm) {
  pivot_longer(varImpFrame.list[[nm]], -one_of('meanImp', 'variable', 'avgRank','avgAbsRank'), 
               names_to = "iteration", values_to = "importance") %>%
    ggplot(.) +
    geom_vline(xintercept = 1, linetype = 2, color = "black") +
    geom_jitter(aes(x = importance, y = variable, color = avgRank), size = 4, alpha = .45, height = .1) + 
    #geom_point(aes(x = meanImp, y = variable), color = "red", size = 4.5, alpha = .05) +
    # stat_summary(aes(x = importance, y = variable),
    #                  fun.data = "mean_cl_boot", colour = "red") +
    stat_summary(aes(x = importance, y = variable), fun = "mean", color = "red") +
    #theme_linedraw() + 
    scale_color_viridis_c() + 
    guides(color = guide_colorbar(reverse = T, title = "Average\nRank", title.hjust = .5)) + 
    theme(panel.grid.major.y = element_blank(), plot.title = element_text(hjust = .5, size = 20), 
          axis.title = element_text(size = 18), axis.text = element_text(size = 15)) +
    labs(title = paste0("Each Variable's Average Permutation Importance \nPredicting ", 
                        nm)) + 
    xlab("%-RMSE Improvement") + ylab("Variable")
  ggsave(paste0("results/regression/2020-06-02/GBM-pviPlot-final_",nm,".jpg"), 
         units = "in", width = 11, height = 8.7, device= "jpg")
})
```



