---
title: "Analysis of Deviation from Structural Brain Volume Centroids"
author: "Ian J. Douglas"
date: "3/2/2020"
output: html_document
---

# Distributions of structural brain volume
Examine the covariance of age, age of adoption,and group classification with predictor features
```{r, results='hide', warning=FALSE, message=FALSE}
if (isFALSE(require(tidyverse))) {install.packages("tidyverse"); library(tidyverse)} else library(tidyverse)
```

### load the master data frame (containing adoption age)
```{r}
# This data will read in with the object name "master_SB" in the environment
load(
  "../../data/raw/3_ALL_long_cort_symptoms_brain_structure_function_epi_ages4-19_2019-05-15.Rdata"
)
# Read in the structural data
strData = readRDS("../../data/processed/structuralLabelled.rds") %>%
# retain only the complete cases (with respect to these structural data; ignoring some labels in filtering)
  filter(., complete.cases(select(.,
    -one_of("IDENT_SUBID","GROUP", "age", "SUBJECTID_long", "wave_to_pull", "cbcl_totprob_t"))))
```

## compute everyone's distance from the structural centroid
Taking into account each structural feature simultaneously, compute the centroid vector of the multivariate cloud of structural features.
```{r}
# Here we compute the centroid of the structural data (for the whole population):
str_unadj_centroid = strData %>% 
  # deselect some labels in the dataset
  select(-one_of("IDENT_SUBID","GROUP", "age", "SUBJECTID_long", "wave_to_pull", "cbcl_totprob_t")) %>%
  summarize_all(~mean(., na.rm = TRUE))

# for each subject, compute the distance between their row vector of the same variables, and the centroid
deviation = rep(NA, times = nrow(strData))
for (i in 1:nrow(strData)) {
  dat = rbind(strData[i, names(str_unadj_centroid)],
              str_unadj_centroid)
  if (all(!is.na(dat))) {
    dist = as.numeric(dist(as.matrix(dat), method = 'euc')) # use euclidean distance
    deviation[i] = dist
  }
}
```

### prepare the data for plotting; plot:
```{r}
plt_dat = cbind(select(strData, GROUP,age), 
                data.frame(dist = deviation))
ggplot(data=plt_dat, aes(x = dist)) +
  geom_histogram(aes(y = ..density.., fill = GROUP), 
                 binwidth=density(na.omit(plt_dat$dist))$bw) +
  ggtitle("Histogram of Euclidean Distances from Population Centroid within each Group") +
  theme(title = element_text(hjust = .5))
```

## Predict deviation from covariates
### Predict deviation from centroid with age, group, and their interaction as covariates
```{r}
# relationship in the aggregate:
summary(lm(dist ~ age * GROUP, data = plt_dat))
```

No significant main effect was found for age, though an interaction and group effect persist.

### Simple main effects within the PI group:
```{r}
summary(lm(dist ~ age, data = plt_dat[plt_dat$GROUP=="PI",]))
```

### Simple main effects within the COMP group:
```{r}
summary(lm(dist ~ age, data = plt_dat[plt_dat$GROUP=="COMP",]))
```

## Now compute centroids within each group, repeat deviation analyses
```{r}
comp_str_unadj_centroid = strData %>%
  filter(GROUP=="COMP") %>%
  select(-one_of("IDENT_SUBID","GROUP", "age", "SUBJECTID_long", "wave_to_pull", "cbcl_totprob_t")) %>%
  summarize_all(~mean(., na.rm = TRUE))

# compute the distances for each comp participant to their group's centroid
comp_deviation = rep(NA, times = nrow(strData))

for (i in grep("COMP", strData$GROUP)) {
  dat = rbind(strData[i, names(comp_str_unadj_centroid)],
              comp_str_unadj_centroid)
  if (all(!is.na(dat))) {
    dist = as.numeric(dist(as.matrix(dat), method = 'euc'))
    comp_deviation[i] = dist
  }
}

# and for PI group:
pi_str_unadj_centroid = strData %>%
  filter(GROUP=="PI") %>%
  select(-one_of("IDENT_SUBID","GROUP", "age", "SUBJECTID_long", "wave_to_pull", "cbcl_totprob_t")) %>%
  summarize_all(~mean(., na.rm = TRUE))

pi_deviation = rep(NA, times = nrow(strData))

for (i in grep("PI", strData$GROUP)) {
  dat = rbind(strData[i, names(pi_str_unadj_centroid)],
              pi_str_unadj_centroid)
  if (all(!is.na(dat))) {
    dist = as.numeric(dist(as.matrix(dat), method = 'euc'))
    pi_deviation[i] = dist
  }
}
```

### Plot the group means, and within-group distances from own-group centroids
```{r, message=FALSE}
barplot_dat = data.frame(Group = c("PI","COMP"), rbind(pi_str_unadj_centroid, comp_str_unadj_centroid))
grp_plt_dat = cbind(
  select(strData, GROUP,age), 
  data.frame(dist = deviation, comp_dist=comp_deviation, pi_dist=pi_deviation)
) %>%
  mutate(withinGroupDeviation = coalesce(comp_dist, pi_dist))

# equivalent to gather(barplot_dat, key="variable", value="value",-Group) %>%
pivot_longer(barplot_dat, cols = -Group, names_to = "variable", values_to = "value") %>%
  ggplot(data = .) +
  geom_bar(aes(x=variable, y = value, fill = Group), stat="identity", position ='dodge') +
  facet_wrap(~variable, scales = 'free') +
  ggtitle("Group means for each structural variable") +
  theme(title = element_text(hjust=.5)) +
  xlab(label = NULL) + ylab(label = NULL) +
  labs(caption = "Generally, PI group have lower mean structural volume, with some exceptions")
#density plot
ggplot(grp_plt_dat, aes(x = withinGroupDeviation)) +
  geom_histogram(aes(y = ..density.. ,fill=GROUP)) +
  ggtitle("Histogram of Euclidean Distances from one's own group's Centroid by Group")  +
  theme(title = element_text(hjust=.5))
```

### Linear model predicting within-group deviation from age
```{r}
summary(lm(withinGroupDeviation ~ age, data = grp_plt_dat))
```

### add interaction term with Group
```{r}
summary(lm(withinGroupDeviation ~ age*GROUP, data = grp_plt_dat))
```

### No main effect, so check simple effects within each group
#### PIs:
```{r}
summary(lm(withinGroupDeviation ~ age, data = grp_plt_dat[grp_plt_dat$GROUP=="PI",]))
```
#### COMPs:
```{r}
summary(lm(withinGroupDeviation ~ age, data = grp_plt_dat[grp_plt_dat$GROUP=="COMP",]))
```


## Now with age of adoption as the IV
### Plots:
```{r}
ageadopt_plt_dat = cbind(
  select(strData, IDENT_SUBID, GROUP, age), 
  data.frame(dist = deviation, comp_dist=comp_deviation, pi_dist=pi_deviation)
) %>%
  mutate(withinGroupDeviation = coalesce(comp_dist, pi_dist)) %>%
  filter(GROUP=="PI") %>% select(-dist:-pi_dist) %>%
  merge(
    .,
    monster_SB %>% filter(!duplicated(IDENT_SUBID)) %>% select(IDENT_SUBID, AGEADOPT),
    by = "IDENT_SUBID", all.y = FALSE
  ) %>%
  # finally, retain only participants who have completed age of adoption data:
  filter(., complete.cases(select(., AGEADOPT)))
# fit the model:
ageadopt_mod = lm(scale(withinGroupDeviation) ~ AGEADOPT, data = ageadopt_plt_dat)

ggplot(data = ageadopt_plt_dat, aes(x=AGEADOPT, y=withinGroupDeviation)) +
  geom_point() +
  #stat_summary(fun.data=mean_cl_normal) +
  geom_smooth(method='lm', formula= y~x) +
  ggtitle("Relationship between age of adoption, and distance from one's own group's centroid",
          subtitle = paste0("slope=",round(ageadopt_mod$coefficients[2], 4)," (n.s.)")) +
  theme(title = element_text(hjust = .5))
```

### Linear model results:
#### Note: the scaled Euclidean distance is predicted.
```{r}
summary(ageadopt_mod)
```


