---
title: "cbclRandomForestRegression"
author: "Ian Douglas"
date: "4/15/2020"
output: html_document
---
# read in the long format cbcl data
```{r}
cbcl <- readRDS("../../../../data/master/final_cbcl_LongFormat.rds") %>%
  # for later merging, recode the group columnn
  mutate_at("GROUP", ~case_when(
    . == 0 ~ "COMP", . == 1 ~ "PI", . == 2 | . == 3 ~ NA_character_
  ))
```

# read in the subcortical brain region variable names
```{r}
subCortVarNames <- read.csv("../../../../data/structuralVarNames20191002_MVT.csv",
                            stringsAsFactors = FALSE) %>%
  filter(X...don.t.include != "X") %>%
  filter(!var %in% c("rhCortexVol", "lhCortexVol")) %>%
  mutate_at('var', ~ifelse(grepl("^CC_", .), ., gsub("_", "\\.", .)))
```

# Read in the brain data; select the above variables
```{r}
load( # loads a data frame called "monster_SB"
  "../../../../data/raw/3_ALL_long_cort_symptoms_brain_structure_function_epi_ages4-19_2019-05-15.Rdata"
)

brainData <- monster_SB %>% 
  select(all_of(c("IDENT_SUBID","index_wave", "GROUP", unlist(subCortVarNames$var)))) %>%
  # right off the bat, delete any wave 3 data
  filter(as.numeric(as.character(index_wave)) < 3) %>%
  # additionally, if any row is missing all brain data, remove the row
  filter(rowSums(select(., -IDENT_SUBID:-GROUP), na.rm = TRUE) != 0) %>%
  # simplify in dex_wave to a numeric variable
  mutate_at("index_wave", ~as.numeric(as.character(.))) 
```

# delete any rows for which the outcome does not exist
```{r}
cbclInt <- cbcl %>% select(IDENT_SUBID:GROUP, CBCL_INTERNPROB_T) %>% drop_na(CBCL_INTERNPROB_T)
cbclExt <- cbcl %>% select(IDENT_SUBID:GROUP, CBCL_EXTERNPROB_T) %>% drop_na(CBCL_EXTERNPROB_T)
cbclTot <- cbcl %>% select(IDENT_SUBID:GROUP, CBCL_TOTPROB_T) %>% drop_na(CBCL_TOTPROB_T)
# also, filter for waves 1 and 2 only.
data.list <- lapply(list(cbclInt, cbclExt, cbclTot), function(x) {
  x %>% filter(as.numeric(as.character(index_wave)) < 3) %>%
    # also, simplift index_wave to a numeric var:
    mutate_at("index_wave", ~as.numeric(as.character(.)))
})
# nnow the data are in a list; attach names correspondingly
names(data.list) <- c("internalizing", "externalizing", "total")
```

# merge the predictor and response data frames
```{r}
# Run an anonymous function to optimally join the data by:
# 1. If the subject has cbcl AND brain data at timepoint 1, use that
# 2. Else, use timepoint 2 data
# (3) If  no match exists, drop the subject
data.list <- lapply(data.list, function(x) {
  # create the union of all the wave 1 data
  wave1 <- merge(x %>% filter(index_wave == 1), 
                 brainData %>% filter(index_wave == 1),
                 by = c("IDENT_SUBID", "index_wave", "GROUP"), all = T)
  # create the union of all the wave 2 data
  wave2 <- merge(x %>% filter(index_wave == 2),
                 brainData %>% filter(index_wave == 2),
                 by = c("IDENT_SUBID", "index_wave", "GROUP"), all = T)
  # append the two
  rbind(wave1, wave2) %>%
    # arrange by subject and index_wave, putting timepoint 1 above timpeoint 2
    arrange(IDENT_SUBID, index_wave) %>%
    # delete the second of any duplicated subjects, if any
    filter(!duplicated(IDENT_SUBID)) # (returned)
})
```

# Adjusting for ICV, sex and age.
#### Here, we also start to pair down the rows, to delete subjects without complete brain data.

First, from the master data frame, read in age and sex, and adjust the structural variables accordingly
```{r}
# make sure the age and gender is correct (wasn't dropped during merging if we have it in the master)
master <- read.csv("../../../../data/master/J.Data_Entry_Master_8_17_17_fixed_fixed_3.csv",
                   stringsAsFactors = F)

data.list2 <- lapply(data.list, function(df) {
  dat <- df # make a copy
  for (i in 1:nrow(dat)) {
    who <- dat$IDENT_SUBID[i]; when <- dat$index_wave[i]
    ageNames <- c("SUBAGE_Session1", "SUBAGE_session2", "SUBAGE_session3")[when]
    # convert the age to years
    dat[i, "age"] <- unlist(master[master$IDENT_SUBID == who, ageNames]) / 12
    dat[i, "gender"] <- ifelse(master$DEM_3_GENDER_CHILD[master$IDENT_SUBID == who] == 1,
                               "Female", "Male")
  }
  dat
})
```
Adjust all structural variables by ICV
```{r}
data.icvAdj <- list()
data.icvAdj$noCovAdj <- lapply(data.list2, function(x) {
  # select columns with complete brain data (excluding estimated total cranial vol for now)
  X <- x %>% filter(rowSums(select(., Left.Thalamus.Proper:CC_Anterior), na.rm = T) != 0)
  mutate_at(X, vars(Left.Thalamus.Proper:CC_Anterior), 
            function(col) {resid(lm(col ~ EstimatedTotalIntraCranialVol, data = X))}) %>%
    # finally, once adjusting for ICV, it can be dropped.
    select(-EstimatedTotalIntraCranialVol)
    
})
# make a comparable list of data frames without adjusted (by ICV) volumes, filtered by complete brain data
data.rawBrain <- list()
data.rawBrain$noCovAdj <- lapply(data.list2, function(x) {
  x %>% filter(rowSums(select(., Left.Thalamus.Proper:CC_Anterior), na.rm = T) != 0)
})
# "data.rawBrain" will contain data frames that can be used with ICV as a predictor, or not
```
Define a function to automate adjusting for any covariate, if it/they significantly covary with a predictor
```{r}
adjustFun <- function(col_name, ..., data)
{
  model.data <- setNames(data[c(col_name, ...)],
                         nm = c("y", ...)) # rename the predictor to "y"
  classes <- sapply(c(...), function(nm) class(model.data[, nm])) # record the covariate type
  # Test for a sig covariance with covariate; based on its class use cor.test or ANOVA:
  sigs <- NULL
  for (i in 1:length(c(...))) {
    if (classes[i] == "factor") {
      f <- as.formula(paste0("y ~ ", c(...)[i]))
      sigs[i] <- summary(lm(formula = f, data = model.data))$coefficients[2, "Pr(>|t|)"] <= .05
    } else sigs[i] <- cor.test(model.data[, "y"], model.data[, c(...)[i]])$p.value <= .05
  }
  # If a relationship is found, extract the residuals from a regression
  if (any(sigs)) {
    sig_covs <- c(...)[sigs]
    f <- as.formula(paste0("y ~ ", paste(sig_covs, collapse = " + ")))
    # now get the residuals:
    resids <- resid(lm(formula = f, data = model.data[, c("y", sig_covs)]))
  } else resids <- data[, col_name] - mean(data[, col_name]) # just mean-center it instead.
  resids # output the residuals (or the original variable if no relationship was found)
}
```
# In this step, we lose any subjets whose age is not known too
```{r}
data.icvAdj$covAdj <- lapply(data.icvAdj$noCovAdj, function(dat) {
  df <- dat %>% drop_na(age, gender)
  cbind.data.frame(df %>% select(IDENT_SUBID:starts_with("CBCL")),
                   setNames(as.data.frame(
                     lapply(names(df %>% select(-IDENT_SUBID:-starts_with("CBCL"))), function(col) {
                       adjustFun(col_name = col, "age", "gender", data = df)
                      })
                   ), nm = names(df %>% select(-IDENT_SUBID:-starts_with("CBCL")))))
})
```
# Also adjust the structural covariate for age and gender that were not adjusted for ICV
```{r}
data.rawBrain$covAdjusted <- lapply(data.rawBrain$noCovAdj, function(dat) {
  df <- dat %>% drop_na(age, gender)
  cbind.data.frame(df %>% select(IDENT_SUBID:starts_with("CBCL")),
                   setNames(as.data.frame(
                     lapply(names(df %>% select(-IDENT_SUBID:-starts_with("CBCL"))), function(col) {
                       adjustFun(col_name = col, "age", "gender", data = df)
                      })
                   ), nm = names(df %>% select(-IDENT_SUBID:-starts_with("CBCL")))))
})
```
# As a final step, delete any rows for whom the response does not exist
```{r}
# data.icvAdj$covAdj <- lapply(data.icvAdj$covAdj, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
# data.icvAdj$noCovAdj <- lapply(data.icvAdj$noCovAdj, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
# data.rawBrain$noCovAdj <- lapply(data.rawBrain$noCovAdj, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
# data.rawBrain$covAdjusted <- lapply(data.rawBrain$covAdjusted, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
```

# Fit models for the ICV-adjusted predictors (with and without adjustement for demographics)
```{r}
# prep the data by dropping NA from the response variable
rfDF.icvAdj <- list()
rfDF.icvAdj$noCovAdj <- lapply(data.icvAdj$noCovAdj, function(d) d %>% drop_na(starts_with("CBCL")))
rfDF.icvAdj$covAdj <- lapply(data.icvAdj$covAdj, function(d) d %>% drop_na(starts_with("CBCL")))
rfDF.rawBrain <- list()
rfDF.rawBrain$noCovAdj <- lapply(data.rawBrain$noCovAdj, function(d) d %>% drop_na(starts_with("CBCL")))
rfDF.rawBrain$covAdjusted <-lapply(data.rawBrain$covAdjusted, 
                                   function(d) d %>% drop_na(starts_with("CBCL")))
# define function to fit the model
# create an option to include or drop the last column (which ICV where it exists)
rf_func <- function(df, icv = FALSE)
{
  model.data <- df %>%
    select(y = starts_with("CBCL"), Left.Thalamus.Proper:last_col(offset = as.numeric(icv)))
  randomForest(y ~ ., data = model.data,
               mtry = 6, ntree = 1000, importance = T, strata = df$GROUP)
}
# Fit the models
icv.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_func)})
raw.withICV.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_func)})
raw.noICV.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, function(x) rf_func(x, icv=TRUE))})
```
# Results
```{r}
rf_model_list <- list(
  "icvAdjusted" = icv.rf_models,
  "notAdjustedByICV" = raw.withICV.rf_models,
  "notICVAdjustedNotInclAsPredictor" = raw.noICV.rf_models
)
saveRDS(rf_model_list, "../../../CBCL_pipeline/output/rf_models_allVars.rds")
lapply(rf_model_list, function(results) {lapply(results, function(list) {lapply(list, function(rf) rf)})})
```

# Re-fit the models on the best variables
```{r}
training.data <- raw_str_csv %>%
  select(SUBJECTID_long, all_of(names(brainData[-1:-3]))) %>%
  filter(!grepl("fu2$", SUBJECTID_long)) %>%
  filter(rowSums(select(., -SUBJECTID_long), na.rm = T)  != 0) %>%
  mutate(IDENT_SUBID = substr(SUBJECTID_long, 1, 5)) %>% select(-SUBJECTID_long) %>%
  filter(!duplicated(IDENT_SUBID)) %>%
  select(IDENT_SUBID, everything()) %>%
  merge(., 
        cbcl %>% 
          select(., IDENT_SUBID, GROUP) %>%
          filter(!duplicated(IDENT_SUBID)) %>%
          filter(!is.na(GROUP)), 
        by = "IDENT_SUBID") %>%
  mutate_at("GROUP", ~factor(., levels = c("PI", "COMP")), all.y = F) %>%
  select(-EstimatedTotalIntraCranialVol)
# fit random forest
rf.train <- randomForest(x = select(training.data, -GROUP, -IDENT_SUBID),
                         y = training.data$GROUP,
                         mtry = 6, ntree = 1000, importance = T, strata = training.data$GROUP,
                         sampsize = rep(sum(training.data$GROUP=="PI"), times = 2))
# permute var imps
pimpRes <- vita::PimpTest(PIMP(X = select(training.data, -GROUP, -IDENT_SUBID), 
                               y = training.data$GROUP,
                               rForest = rf.train,
                               S = 1000))
saveRDS(pimpRes, "../../../CBCL_pipeline/output/pimpResults_structure.rds")
data.frame(pimpRes$pvalue, "var"=rownames(pimpRes$pvalue), stringsAsFactors = FALSE) %>% 
  arrange(p.value) %>% .[1:11,"var"]
```

refit the models with the top 11 of 21 variables
```{r}
bestVars <- data.frame(pimpRes$pvalue, "var"=rownames(pimpRes$pvalue), stringsAsFactors = FALSE) %>% 
  arrange(p.value) %>% .[1:11,"var"]
rf_funcBest <- function(df, icv = FALSE)
{
  model.data <- df %>% select(y = starts_with("CBCL"), all_of(bestVars))
  randomForest(y ~ ., data = model.data,
               mtry = 8, ntree = 1000, importance = T, strata = df$GROUP)
} 
bestVar.icv.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_funcBest)})
bestVar.raw.withICV.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_funcBest)})
```
```{r}
bestVars_rf_model_list <- list(
  "icvAdjusted" = bestVar.icv.rf_models,
  "notAdjustedByICV" = bestVar.raw.withICV.rf_models
)
saveRDS(bestVars_rf_model_list, "../../../CBCL_pipeline/output/bestVars_rf_model_list.rds")
lapply(bestVars_rf_model_list, function(list) {lapply(list, function(rf) rf)})
```

# Variable importance plots
```{r}
source("../../../misc/vimPlot.R")
rf_model_list <- readRDS("../../../CBCL_pipeline/output/rf_models_allVars.rds")
lapply(rf_model_list, function(list) {lapply(list, function(list2) {lapply(list2, vimPlot)})})
```

