---
title: "Structural Data: Importance-Based Variable Selection"
author: "Ian Douglas"
date: "4/8/2020"
output: html_document
---
### Loading packages
```{r, results='hide', warning=FALSE, message=FALSE}
library(tidyverse)
```
### source helper function
```{r}
source("../../Scripts/misc/descending_rank.R")
source("../../Scripts/misc/min_max.R")
```

# Subset set of segmented structural features of non-overlapping regions
### Read in:
    1. The repeated test-set AUC-based VI cross validation with the structural volume predictors
    2. The corresponding dataset on which variable ranking was conducted
```{r, collapse=TRUE}
classificationResults_segSubset <- readRDS(
  "../../output/featureExtraction/segmentationSubset_testSetVarImp_results.rds")
adjStr_segSubset <- readRDS("../../data/processed/adjustedStructural_segmentationSubset_2020-04-08.rds")
```

## Extract OOB AUC scores
There is only one per seed (forest) because this was generated using the predict() function on the fully built forest
```{r}
OOB_AUCs_segSubset <- sapply(classificationResults_segSubset, function(seed) seed$oob.Score)
```
## Extract the test-set AUCs by aggregating each tree's test-set AUC
This requires aggregating because the test-set AUC was computed seperatly at each tree,  to the same test set.
```{r}
crossValTest_AUCs_perSeed_segSubset <- sapply(
  X = classificationResults_segSubset,
  FUN = function(seed) {
    bagged_testsetAUC = mean(sapply(seed$forest, function(tree) {
      tree$auc # this is test-set AUC
    }))
  bagged_testsetAUC
})
```
## Quick plot of the distribution of out-of-bag and test-set AUCs
```{r, fig.asp=.38, fig.width=16}
ggplot(rbind(data.frame("source" = "OOB", "AUC" = OOB_AUCs_segSubset),
             data.frame("source" = "CV", AUC = crossValTest_AUCs_perSeed_segSubset))) + 
  geom_density(aes(x = AUC, fill = source), alpha = .65) +
  theme_linedraw()
```

## Variable Ranking
    1. Within each forest, rank the variables by the absolute value of their importances
    2. Average these ranks across all seeds
```{r}
### First extract the variable importances for each variable, within each seed/forest
impFrame_wide_segSubset <- reduce(
  map(classificationResults_segSubset, function(seed) {
    imps = sapply(seed$testSetImportance, function(var) var$AUC_VIM)
    data.frame("var" = names(seed$testSetImportance), "importance" = imps)
    }),
  left_join, by = "var"
)
names(impFrame_wide_segSubset)[-1] <- paste0("importance_", grep("imp",names(impFrame_wide_segSubset[-1])))
### Compute the repeated cross-validation importance by averaging over all seeds
impFrame_wide_segSubset$avgTestSetImp <- rowMeans(
  impFrame_wide_segSubset[grep("imp", names(impFrame_wide_segSubset))]
)
### Compute the absolute ranks,and average absolute importance
impFrame_wide_segSubset$avgAbsImportance <- rowMeans(
  mutate_all(impFrame_wide_segSubset[grep("imp", names(impFrame_wide_segSubset))], abs)
)
impFrame_wide_segSubset$avgAbsRank <- rowMeans(
  mutate_all(impFrame_wide_segSubset[grep("^imp", names(impFrame_wide_segSubset))],
             ~descending_rank(abs(.)))
)

## sort the data frame for plotting purposes
impFrame_wide_segSubset <- impFrame_wide_segSubset %>%
  arrange(desc(avgAbsImportance)) %>%
  mutate_at("var", ~factor(., levels = rev(.)))
```

## View the variable importances
```{r}
ggplot(impFrame_wide_segSubset) +
  geom_col(aes(x = var, y = avgAbsImportance, fill = avgAbsRank)) +
  scale_fill_gradient(low = "#00906f", high = "#ffbc00", 
                      guide = guide_colorbar(reverse = TRUE)) +
  coord_flip()
```

# Repeat, with the set of all features extracted from the structural MRI scans
### Read in:
1. The repeated test-set AUC-based VI cross validation with the structural volume predictors
2. The corresponding dataset on which variable ranking was conducted
```{r}
classificationResults <- readRDS(
  "../../output/featureExtraction/NEWEST_adjSTR_testSetAUC_VIM_results.rds")
adjStr <- readRDS("../../data/processed/adjustedStructuralData_2020-04-08.rds")
```

## Extract OOB AUC scores
There is only one per seed (forest) because this was generated using the predict() function on the fully built forest
```{r}
OOB_AUCs <- sapply(classificationResults, function(seed) seed$oob.Score)
```
## Extract the test-set AUCs by aggregating each tree's test-set AUC
This requires aggregating because the test-set AUC was computed seperatly at each tree,  to the same test set.
```{r}
crossValTest_AUCs_perSeed <- sapply(
  X = classificationResults,
  FUN = function(seed) {
    bagged_testsetAUC = mean(sapply(seed$forest, function(tree) {
      tree$auc # this is test-set AUC
    }))
  bagged_testsetAUC
})
```
## Distribution of test-set and OOB AUCs:
```{r, fig.asp=.38, fig.width=16}
ggplot(rbind(data.frame("source" = "OOB", "AUC" = OOB_AUCs),
             data.frame("source" = "CV", AUC = crossValTest_AUCs_perSeed))) + 
  geom_density(aes(x = AUC, fill = source), alpha = .7) +
  theme_linedraw()
```

## Variable Ranking
    1. Within each forest, rank the variables by the absolute value of their importances
    2. Average these ranks across all seeds
```{r}
### First extract the variable importances for each variable, within each seed/forest
impFrame_wide <- reduce(
  map(classificationResults, function(seed) {
    imps = sapply(seed$testSetImportance, function(var) var$AUC_VIM)
    data.frame("var" = names(seed$testSetImportance), "importance" = imps)
    }),
  left_join, by = "var"
)
names(impFrame_wide)[-1] <- paste0("importance_", grep("imp",names(impFrame_wide[-1])))
### Compute the repeated cross-validation importance by averaging over all seeds
impFrame_wide$avgTestSetImp <- rowMeans(
  impFrame_wide[grep("imp", names(impFrame_wide))]
)
### Compute the absolute ranks,and average absolute importance
impFrame_wide$avgAbsImportance <- rowMeans(
  mutate_all(impFrame_wide[grep("imp", names(impFrame_wide))], abs)
)
impFrame_wide$avgAbsRank <- rowMeans(
  mutate_all(impFrame_wide[grep("^imp", names(impFrame_wide))],
             ~descending_rank(abs(.)))
)

## sort the data frame for plotting purposes
impFrame_wide <- impFrame_wide %>%
  arrange(desc(avgAbsImportance)) %>%
  mutate_at("var", ~factor(., levels = rev(.)))
```

## View the variable importances
```{r, fig.asp=.6, fig.width=16}
ggplot(impFrame_wide) +
  geom_col(aes(x = var, y = avgAbsImportance, fill = avgAbsRank)) +
  scale_fill_gradient(low = "#00906f", high = "#ffbc00", 
                      guide = guide_colorbar(reverse = TRUE)) +
  coord_flip()
```


