---
title: "Model-Based Recursive Partitioning"
author: "Ian Douglas"
date: "4/15/2020"
output: html_document
---
# read in the long format cbcl data
```{r}
# cbcl <- readRDS("../../../../data/master/final_cbcl_LongFormat.rds") %>%
#   # for later merging, recode the group columnn
#   mutate_at("GROUP", ~case_when(
#     . == 0 ~ "COMP", . == 1 ~ "PI", . == 2 | . == 3 ~ NA_character_
#   ))
```
### required packages
```{r, message=FALSE, results='hide', warning=FALSE}
library(tidyverse)
library(haven)
library(MLmetrics)
library(parallel)
library(partykit)
library(randomForest)
```
# load in the SB master data frame with the CBCL collected (if and) when the participants were older
```{r}
sb.master <- haven::read_sav("../../data/master/J.Data_Entry_Master_8_17_17_fixed_fixed_4.sav")
```
# read in the subcortical brain predictors (adjusted and unadjusted)
```{r}
labelledBrainData <- readRDS("../../data/master/masterStrDataLong.rds") # unadjusted
adjustedData <- readRDS("../../data/master/masterAdjustedStrDataLong_2020-04-28.rds") # adjusted
```
# select the desired cbcl data from the master data frame.
```{r}
cbcl <- sb.master %>% # from the data frame in which the cbcl survey is the older kids
  select(IDENT_SUBID, contains("SUBAGE_s", ignore.case = F), contains("CBCL_")) %>%
  select(IDENT_SUBID, contains("SUBAGE_s"), contains("INTERN"), contains("EXTERN"), contains("TOTPROB")) %>%
  select(IDENT_SUBID, contains("SUBAGE_s"), contains("PROB_T"), contains("PROB_TOT")) %>%
  # reorder the columns
  select(1:4, matches("_T$"), matches("TOT$"), contains("FU1"), contains("FU2"), everything()) %>%
  # extend the ages down into long form & use them to create a WAVE column to match that in the brain data
  gather(., key = 'WAVE', value = 'brain_age_yrs', SUBAGE_session2:SUBAGE_session4) %>%
  # convert it to yrs
  mutate_at('brain_age_yrs', ~ ./12) %>%
  select(IDENT_SUBID, WAVE, brain_age_yrs, everything()) %>% # reorder the columns again
  # now change the subage_session2 to WAVE2, and so forth for session3 and 4; so, (subage_session2 --> WAVE1)
  mutate_at("WAVE", ~paste0("WAVE", as.numeric(sub("SUBAGE_session", "", .)) - 1)) %>%
  arrange(IDENT_SUBID, WAVE)
```
# Collapse the entries for each timepoint from the cbcl_1.5 and cbcl_4 for each category
```{r}
INTRNLZNG_T = NULL; INTRNLZNG_TOT = NULL
EXTRNLZNG_T = NULL; EXTRNLZNG_TOT = NULL
TOTALPROB_T = NULL; TOTALPROB_TOT = NULL
# define quick helper func
get.dat <- function(pat1, pat2)
{
  na.omit(unlist(
    cbclDat[i, grepl(pat1, names(cbclDat)) & grepl(pat2, names(cbclDat))]
  ))
}

for (i in 1:nrow(cbcl)) {
  if (cbcl$WAVE[i] == "WAVE1") {
    cbclDat <- cbcl %>% select(ends_with("T")) # either "T" or "TOT" desired
  } else if (cbcl$WAVE[i] == "WAVE2") {
    cbclDat <- cbcl %>% select(ends_with("_FU1"))
  } else cbclDat <- cbcl %>% select(ends_with("_FU2"))
  
  res <- get.dat("INTERNPROB", "_T$|_T_")
  INTRNLZNG_T[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("INTERNPROB", "_TOT$|_TOT_")
  INTRNLZNG_TOT[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("EXTERNPROB", "_T$|_T_")
  EXTRNLZNG_T[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("EXTERNPROB", "_TOT$|_TOT_")
  EXTRNLZNG_TOT[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("TOTPROB", "_T$|_T_")
  TOTALPROB_T[i] <- ifelse(length(res) == 1, res, NA)
  res <- get.dat("TOTPROB", "_TOT$|_TOT_")
  TOTALPROB_TOT[i] <- ifelse(length(res) == 1, res, NA)
}
```
# Compile from the above results the cbcl long format data
```{r}
cbcl_long = data.frame(
  select(cbcl, IDENT_SUBID, WAVE, brain_age_yrs),
  CBCL_INTERNALIZING_T = INTRNLZNG_T, CBCL_INTERNALIZING_TOT = INTRNLZNG_TOT,
  CBCL_EXTERNALIZING_T = EXTRNLZNG_T, CBCL_EXTERNALIZING_TOT = EXTRNLZNG_TOT,
  CBCL_TOTALPROB_T = TOTALPROB_T, CBCL_TOTALPROB_TOT = TOTALPROB_TOT,
  stringsAsFactors = F
)
```

# merge in the data frames together
```{r, message=FALSE}
cbclAdjBrain <- left_join(x = adjustedData, y = cbcl_long,
                          by = c("IDENT_SUBID", "WAVE", "brain_age_yrs"),
                          all.x = T, all.y = F) %>%
  select(IDENT_SUBID:WAVE, subjWeight, everything()) %>% ungroup()  %>%
  # as a final step, reorder the levels of the GROUP factor (PI first, then COMP)
  mutate_at("GROUP", ~factor(., levels = c("PI", "COMP")))
# same for the unadjusted brain data:
cbclRawBrain <- left_join(x = labelledBrainData, y = cbcl_long,
                          by = c("IDENT_SUBID", "WAVE", "brain_age_yrs"),
                          all.x = T, all.y = F) %>%
  group_by(IDENT_SUBID) %>%
  mutate(subjWeight = 1 / n()) %>%
  select(IDENT_SUBID:WAVE, subjWeight, everything()) %>% ungroup()  %>%
  # as a final step, reorder the levels of the GROUP factor (PI first, then COMP)
  mutate_at("GROUP", ~factor(., levels = c("PI", "COMP")))
```

# Fit the MOBs
```{r}
reg <- function(y, x, start = NULL, weights = NULL, offset = NULL, ...) 
{ 
  glm(y ~ 0 + x, family = gaussian, start = start, ...)
}
```
```{r}
f1 <- as.formula(paste0("CBCL_INTERNALIZING_T ~ GROUP | ", regressors))
# consider using glmertree
mob1 <- partykit::mob(formula = f1,  data = cbclAdjBrain, 
                      weights = adjustedData$subjWeight, na.action = na.omit, fit = reg,
                      control= mob_control(alpha = 1)) # skip the sig test to control splits
mob1
```
```{r}
plot(mob1)
```
```{r}
AIC(partykit::mob(formula = f1,  data = cbclAdjBrain, 
                      weights = adjustedData$subjWeight, na.action = na.omit, fit = reg,
                      control= mob_control(alpha = 1, minsize = 1, prune = "AIC")))
```

```{r}
mob2 <- partykit::mob(formula = f1,  data = cbclRawBrain, 
                      weights = cbclRawBrain$subjWeight, na.action = na.omit, fit = reg,
                      control= mob_control(alpha = 1))
summary(mob2)
plot(mob2)
```
```{r}
f3 <- as.formula(paste0("CBCL_INTERNALIZING_T ~ GROUP+brain_age_yrs+GENDER_FEMALE | ", regressors))
partykit::mob(f3,  data = cbclRawBrain, 
              weights = cbclRawBrain$subjWeight, na.action = na.omit, fit = reg)
```
```{r}
f4 <- as.formula(
  paste0("CBCL_INTERNALIZING_T ~ GROUP+",covariates,"|",regressors)
)
mob(f4,  data = cbclRawBrain, 
              weights = cbclRawBrain$subjWeight, na.action = na.omit, fit = reg)
```
```{r}
f5 <- as.formula(
  paste0("CBCL_INTERNALIZING_T ~ -1+",regressors,"| GROUP")
)
mob(f5,  data = cbclRawBrain, 
    weights = cbclRawBrain$subjWeight, na.action = na.omit, fit = reg)
```

# Select the earlier data, and predict cbcl in RF
```{r}
rf_data.frames <- lapply(list(cbclAdjBrain, cbclRawBrain), function(df) {
  df.list <- lapply(grep("CBCL",names(df),value = T), function(var.nm) {
    df %>%
      select(1:(grep("CBCL",names(.))[1]-1), all_of(var.nm)) %>%
      drop_na(!!rlang::enquo(var.nm)) %>% # drop_na with respect to the cbcl column
      # finallly, if duplicates still persist, select the earlier instance
      filter(!duplicated(IDENT_SUBID))
  })
  names(df.list) <- grep("CBCL",names(df),value = T)
  df.list
})
names(rf_data.frames) <- c("adjusted","raw")
lapply(rf_data.frames, function(rfdf) {lapply(rfdf, dim)})
```

# for each outcome run the rf function
```{r}
# Define functions to do the whole process on all data frames in the above-created list.

# function to uniformly preprocess all data frames identically:
preProcRegular <- function(my_data)
{
  my_data %>% 
    rename_at(vars(contains("CBCL")), ~replace(., TRUE, "y")) %>%
    select(-GENDER_FEMALE, -brain_age_yrs, -WAVE, -subjWeight, -EstimatedTotalIntraCranialVol) %>%
    column_to_rownames("IDENT_SUBID")
}
# function to uniformly fit the rf the same way every time:
fit_RF <- function(.data, .preProc = NULL)
{
  if (is_null(.preProc)) {
    the_data <- .data
  } else the_data <- .preProc(.data)
  randomForest(
    y ~ .-GROUP,
    data = the_data,
    strata = the_data$GROUP,
    sampsize = sum(the_data$GROUP=="PI") * n_distinct(the_data$GROUP),
    mtry = sqrt(ncol(the_data) - 8),
    ntree = 901, importance = T
  )
}
# function to wrap around the above to fit RF, and then compute LOOCV on it:
fit_and_score <- function(data, fit_func, cv = "loocv", preProc = NULL)
{
  if (cv != "loocv") stop("loocv only for now")
  # The model:
  baseline <- fit_func(data, .preProc = preProc) # (preProc runs inside this func)
  # Repeat for each row heldout:
  LOOCV_scores <- sapply(1:nrow(data), function(i, preproc = preProc) {
    train <- data[-i, ]
    if (is_null(preproc)) {
      test <- data[i, ]
    } else test <- preProc(data[i, ]) # use preProc on this one out here for test.
    fit <- fit_func(train) # preProc happens inside here for train.
    MLmetrics::RMSE(
      y_pred = predict(fit, newdata = test),
      y_true = test[, ncol(test)] # recall that CBCL is in the last column
    )
  })
  # return a named list
  nms <- c(paste0("RF.",names(select(data, last_col()))),"LOOCV_scores","LOOCV_mean")
  out <- list(baseline, LOOCV_scores, mean(LOOCV_scores))
  names(out) <- nms
  out
}

# fit this setup on all data frames in the nested list
rf_models_earlierWave <- lapply(rf_data.frames, function(rfdfs) {
  mclapply(X = rfdfs, FUN = function(df) {
    fit_and_score(data = df, fit_func = fit_RF, cv = "loocv", preProc = preProcRegular)
  }, mc.cores = 3)
})
```
```{r}
# save results
saveRDS(rf_models_earlierWave, "../CBCL_pipeline/output/earlierWaveCBCL-RF_allModels_2020-05-04.rds")
```

# Print results
```{r}
lapply(rf_models_earlierWave, function(list) {
  lapply(list, function(mod) mod[[1]])
})
```
```{r}
lapply(rf_models_earlierWave, function(list) {
  lapply(list, function(mod) varImpPlot(mod[[1]],type=1))
})
```

# Repeat the above, but regress GROUP out of the outcome variable
```{r}
# Define a new preProc function and repeat.

# function to uniformly preprocess all data frames identically:
preProcGrpAdjust <- function(my_data)
{
  out.dat <- my_data %>% 
    rename_at(vars(contains("CBCL")), ~replace(., TRUE, "y")) %>%
    select(-GENDER_FEMALE, -brain_age_yrs, -WAVE, -subjWeight, -EstimatedTotalIntraCranialVol) %>%
    column_to_rownames("IDENT_SUBID")
  out.dat$y <- resid(lm(y ~ GROUP, data = out.dat))
  out.dat
}

# fit this setup on all data frames in the nested list
rf_models_earlierWaveGrpCTRL <- lapply(rf_data.frames, function(rfdfs) {
  mclapply(X = rfdfs, FUN = function(df) {
    newdf <- preProcGrpAdjust(df) # do the preproc out here
    fit_and_score(data = newdf, fit_func = fit_RF, cv = "loocv", preProc = NULL)
  }, mc.cores = 3)
})
```
# Print results
```{r}
lapply(rf_models_earlierWaveGrpCTRL, function(L) {lapply(L, function(mod) mod[[1]])})
```

# Regress group out of the predictors
```{r}
# Define a new preProc function to control for GROUP in the predictors. Re-fit
# Function to uniformly preprocess all data frames identically:
preProcGrpCOVadjust <- function(my_data)
{
  out.dat <- my_data %>% 
    rename_at(vars(contains("CBCL")), ~replace(., TRUE, "y")) %>%
    select(-GENDER_FEMALE, -brain_age_yrs, -WAVE, -subjWeight, -EstimatedTotalIntraCranialVol) %>%
    column_to_rownames("IDENT_SUBID") %>%
    # now adjust each brain-data column
    mutate_at(vars(-GROUP, -y),
              function(col) {resid(lm(covar ~ GROUP, data = data.frame(covar = col, GROUP=.$GROUP)))})
  out.dat
}

# fit this setup on all data frames in the nested list
rf_models_earlyWaveGrpCOVadj <- lapply(rf_data.frames, function(rfdfs) {
  mclapply(X = rfdfs, FUN = function(df) {
    newdf <- preProcGrpCOVadjust(df) # do the preproc out here
    fit_and_score(data = newdf, fit_func = fit_RF, cv = "loocv", preProc = NULL)
  }, mc.cores = 3)
})
```
## Results
```{r}
lapply(rf_models_earlyWaveGrpCOVadj, function(L) {lapply(L, function(mod) mod[[1]])})
```

# Baseline models within each group
(NONE WORKED)
```{r}
# lapply(rf_data.frames, function(L) {lapply(L, function(df) {
#   models<-lapply(c("PI","COMP"), function(g) {
#     fit_RF(df[df$GROUP == g, ], .preProc = preProcRegular)
#   })
#   names(models) <- c("PI", "COMP")
#   models
# })})
```

# Using all of the data

Augment the data so that bootstrap resampling results in equal probability of selecting each subject, but is still completely random with respect to which wave is selected.
```{r}
# Note, just conducting this analysis on the adjusted data, and the T score outcomes
augment.rf_data <- lapply(
  X = list(CBCL_INTERNALIZING_T="CBCL_INTERNALIZING_T", 
           CBCL_EXTERNALIZING_T="CBCL_EXTERNALIZING_T", 
           CBCL_TOTALPROB_T="CBCL_TOTALPROB_T"),
  FUN = function(nm) {
    rf_data.frames$adjusted[[nm]] %>%
      rbind(., filter(., subjWeight == 1)) %>%
      select(-GENDER_FEMALE,-brain_age_yrs, -WAVE, -subjWeight, -EstimatedTotalIntraCranialVol) %>%
      # rename the outcome y for uniformity:
      select(IDENT_SUBID, GROUP, y = last_col(), everything())
})

# define a function to produce the trees
rfRegressor <- function(seed, .data)
{
  set.seed(seed)
  randomForest(
    y ~.-GROUP-IDENT_SUBID, data = .data,
    mtry = sqrt(ncol(.data) - 3), # minus GROUP, IDENT_SUBID and y
    ntree = 1, # one tree
    strata = .data$GROUP,
    sampsize = n_distinct(.data$IDENT_SUBID[adjustedData$GROUP=="PI"])*n_distinct(.data$GROUP),
    importance = TRUE,
    keep.inbag = TRUE,
    keep.forest = TRUE,
    localImp = TRUE
  )
}
# Fit a forest of these bootstrapped trees for each outcome
augmented.forest.list <- lapply(augment.rf_data, function(df) {
  mclapply(1:901, function(seed) {
    rfRegressor(seed = seed, .data = df)
  }, mc.cores = 3)
})

# Package up the forests with their original data
aug.RF_pkg <- list(
  CBCL_INTERNALIZING_T=list("df" = augment.rf_data$CBCL_INTERNALIZING_T,
                            "forest"= augmented.forest.list$CBCL_INTERNALIZING_T), 
  CBCL_EXTERNALIZING_T=list("df" = augment.rf_data$CBCL_EXTERNALIZING_T,
                            "forest"= augmented.forest.list$CBCL_EXTERNALIZING_T), 
  CBCL_TOTALPROB_T=list("df"=augment.rf_data$CBCL_TOTALPROB_T,
                        "forest" = augmented.forest.list$CBCL_TOTALPROB_T)
)
# now add to each result the IDENT_SUBID of all subjects in bag and oob
# Note, this is done manually b/c inbag does not mean a subject is not ALSO out of bag
for (i in 1:length(aug.RF_pkg)) {
  for (j in 1:length(aug.RF_pkg[[i]]$forest)) {
    # Figure out who is OOB
    the_model <- aug.RF_pkg[[i]]$forest[[j]]
    inbag.index <- row(the_model$inbag)[, 1][the_model$inbag[, 1] != 0]
    aug.RF_pkg[[i]]$forest[[j]]$INBAG.ID <- aug.RF_pkg[[i]]$df$IDENT_SUBID[inbag.index]
    aug.RF_pkg[[i]]$forest[[j]]$OOB.ID <- aug.RF_pkg[[i]]$df$IDENT_SUBID[ # filter for OOB
      !aug.RF_pkg[[i]]$df$IDENT_SUBID %in% aug.RF_pkg[[i]]$forest[[j]]$INBAG.ID]
    aug.RF_pkg[[i]]$forest[[j]]$the_tree <- getTree(the_model, k = 1, labelVar = TRUE) # also add the tree
    # Now get the out of bag predictions to the people in "OOB.ID"
    aug.RF_pkg[[i]]$forest[[j]]$OOB.PREDICTIONS <- predict(
      the_model, 
      newdata = aug.RF_pkg[[i]]$df[
        aug.RF_pkg[[i]]$df$IDENT_SUBID %in% aug.RF_pkg[[i]]$forest[[j]]$OOB.ID, ]
    )
    # True y:
    YTRUE <- aug.RF_pkg[[i]]$df$y[
      aug.RF_pkg[[i]]$df$IDENT_SUBID %in% aug.RF_pkg[[i]]$forest[[j]]$OOB.ID]
    # RMSE:
    aug.RF_pkg[[i]]$forest[[j]]$OOB.RMSE <- RMSE(
      y_pred = aug.RF_pkg[[i]]$forest[[j]]$OOB.PREDICTIONS,
      y_true = aug.RF_pkg[[i]]$df$y[
        aug.RF_pkg[[i]]$df$IDENT_SUBID %in% aug.RF_pkg[[i]]$forest[[j]]$OOB.ID]
    )
    # Pseudo rsq:
    aug.RF_pkg[[i]]$forest[[j]]$OOB.RSQ <- 1 - ((aug.RF_pkg[[i]]$forest[[j]]$OOB.RMSE^2)/var(YTRUE))
  }
}


# Forest score (aggregation)
for (i in 1:length(aug.RF_pkg)) {
  aug.RF_pkg[[i]]$aggregate_RMSE <- mean(sapply(aug.RF_pkg[[i]]$forest, function(tree) tree$OOB.RMSE))
  aug.RF_pkg[[i]]$aggregate_rsq <- mean(sapply(aug.RF_pkg[[i]]$forest, function(tree) tree$OOB.RSQ), na.rm = T)
}
```


# Another version with older wave data
```{r}
# Select the earlier data, and predict cbcl in RF
rf_Older_data.frames <- lapply(list(cbclAdjBrain, cbclRawBrain), function(df) {
  df.list <- lapply(grep("CBCL",names(df),value = T), function(var.nm) {
    df %>%
      select(1:(grep("CBCL",names(.))[1]-1), all_of(var.nm)) %>%
      drop_na(!!rlang::enquo(var.nm)) %>% # drop_na with respect to the cbcl column
      # finally, select the latest instance of each participant
      group_by(IDENT_SUBID) %>%
      mutate(subjCount = 1:n()) %>%
      filter(subjCount == max(subjCount)) %>%
      select(-subjCount) %>%
      ungroup()
  })
  names(df.list) <- grep("CBCL",names(df),value = T)
  df.list
})
names(rf_Older_data.frames) <- c("adjusted","raw")
lapply(rf_Older_data.frames, function(rfdf) {lapply(rfdf, dim)})
```
## Fit models as above
```{r}
rf_models_olderWave <- lapply(rf_Older_data.frames, function(rfdfs) {
  mclapply(X = rfdfs, FUN = function(df) {
    newdf <- preProcRegular(df)
    fit_and_score(data = newdf, fit_func = fit_RF, cv = "loocv", preProc = NULL)
  }, mc.cores = 3)
})
```
### Results
```{r}
lapply(rf_models_olderWave, function(L) {lapply(L, function(mod) mod[[1]])})
```
# Save this data for further tunin
```{r}
final.df.list <- list()
final.df.list$CBCL_INTERNALIZING_T <- rf_Older_data.frames$adjusted$CBCL_INTERNALIZING_T
final.df.list$CBCL_EXTERNALIZING_T <- rf_Older_data.frames$adjusted$CBCL_EXTERNALIZING_T
final.df.list$CBCL_TOTALPROB_T <- rf_Older_data.frames$adjusted$CBCL_TOTALPROB_T
saveRDS(final.df.list,
        "~/DANL/SB/ianMasters/data/processed/finalStrAdjDF-LIST_2020-05-07.rds")
fit_RF
```


ARCHIVED OLD VERSION
<!---
# Read in the brain data; select the above variables
```{r}
load( # loads a data frame called "monster_SB"
  "../../../../data/raw/3_ALL_long_cort_symptoms_brain_structure_function_epi_ages4-19_2019-05-15.Rdata"
)

brainData <- monster_SB %>% 
  select(all_of(c("IDENT_SUBID","index_wave", "GROUP", unlist(subCortVarNames$var)))) %>%
  # right off the bat, delete any wave 3 data
  filter(as.numeric(as.character(index_wave)) < 3) %>%
  # additionally, if any row is missing all brain data, remove the row
  filter(rowSums(select(., -IDENT_SUBID:-GROUP), na.rm = TRUE) != 0) %>%
  # simplify in dex_wave to a numeric variable
  mutate_at("index_wave", ~as.numeric(as.character(.))) 
```

# delete any rows for which the outcome does not exist
```{r}
cbclInt <- cbcl %>% select(IDENT_SUBID:GROUP, CBCL_INTERNPROB_T) %>% drop_na(CBCL_INTERNPROB_T)
cbclExt <- cbcl %>% select(IDENT_SUBID:GROUP, CBCL_EXTERNPROB_T) %>% drop_na(CBCL_EXTERNPROB_T)
cbclTot <- cbcl %>% select(IDENT_SUBID:GROUP, CBCL_TOTPROB_T) %>% drop_na(CBCL_TOTPROB_T)
# also, filter for waves 1 and 2 only.
data.list <- lapply(list(cbclInt, cbclExt, cbclTot), function(x) {
  x %>% filter(as.numeric(as.character(index_wave)) < 3) %>%
    # also, simplift index_wave to a numeric var:
    mutate_at("index_wave", ~as.numeric(as.character(.)))
})
# nnow the data are in a list; attach names correspondingly
names(data.list) <- c("internalizing", "externalizing", "total")
```

# merge the predictor and response data frames
```{r}
# Run an anonymous function to optimally join the data by:
# 1. If the subject has cbcl AND brain data at timepoint 1, use that
# 2. Else, use timepoint 2 data
# (3) If  no match exists, drop the subject
data.list <- lapply(data.list, function(x) {
  # create the union of all the wave 1 data
  wave1 <- merge(x %>% filter(index_wave == 1), 
                 brainData %>% filter(index_wave == 1),
                 by = c("IDENT_SUBID", "index_wave", "GROUP"), all = T)
  # create the union of all the wave 2 data
  wave2 <- merge(x %>% filter(index_wave == 2),
                 brainData %>% filter(index_wave == 2),
                 by = c("IDENT_SUBID", "index_wave", "GROUP"), all = T)
  # append the two
  rbind(wave1, wave2) %>%
    # arrange by subject and index_wave, putting timepoint 1 above timpeoint 2
    arrange(IDENT_SUBID, index_wave) %>%
    # delete the second of any duplicated subjects, if any
    filter(!duplicated(IDENT_SUBID)) # (returned)
})
```

# Adjusting for ICV, sex and age.
#### Here, we also start to pair down the rows, to delete subjects without complete brain data.

First, from the master data frame, read in age and sex, and adjust the structural variables accordingly
```{r}
# make sure the age and gender is correct (wasn't dropped during merging if we have it in the master)
master <- read.csv("../../../../data/master/J.Data_Entry_Master_8_17_17_fixed_fixed_3.csv",
                   stringsAsFactors = F)

data.list2 <- lapply(data.list, function(df) {
  dat <- df # make a copy
  for (i in 1:nrow(dat)) {
    who <- dat$IDENT_SUBID[i]; when <- dat$index_wave[i]
    ageNames <- c("SUBAGE_Session1", "SUBAGE_session2", "SUBAGE_session3")[when]
    # convert the age to years
    dat[i, "age"] <- unlist(master[master$IDENT_SUBID == who, ageNames]) / 12
    dat[i, "gender"] <- ifelse(master$DEM_3_GENDER_CHILD[master$IDENT_SUBID == who] == 1,
                               "Female", "Male")
  }
  dat
})
```
Adjust all structural variables by ICV
```{r}
data.icvAdj <- list()
data.icvAdj$noCovAdj <- lapply(data.list2, function(x) {
  # select columns with complete brain data (excluding estimated total cranial vol for now)
  X <- x %>% filter(rowSums(select(., Left.Thalamus.Proper:CC_Anterior), na.rm = T) != 0)
  mutate_at(X, vars(Left.Thalamus.Proper:CC_Anterior), 
            function(col) {resid(lm(col ~ EstimatedTotalIntraCranialVol, data = X))}) %>%
    # finally, once adjusting for ICV, it can be dropped.
    select(-EstimatedTotalIntraCranialVol)
    
})
# make a comparable list of data frames without adjusted (by ICV) volumes, filtered by complete brain data
data.rawBrain <- list()
data.rawBrain$noCovAdj <- lapply(data.list2, function(x) {
  x %>% filter(rowSums(select(., Left.Thalamus.Proper:CC_Anterior), na.rm = T) != 0)
})
# "data.rawBrain" will contain data frames that can be used with ICV as a predictor, or not
```
Define a function to automate adjusting for any covariate, if it/they significantly covary with a predictor
```{r}
adjustFun <- function(col_name, ..., data)
{
  model.data <- setNames(data[c(col_name, ...)],
                         nm = c("y", ...)) # rename the predictor to "y"
  classes <- sapply(c(...), function(nm) class(model.data[, nm])) # record the covariate type
  # Test for a sig covariance with covariate; based on its class use cor.test or ANOVA:
  sigs <- NULL
  for (i in 1:length(c(...))) {
    if (classes[i] == "factor") {
      f <- as.formula(paste0("y ~ ", c(...)[i]))
      sigs[i] <- summary(lm(formula = f, data = model.data))$coefficients[2, "Pr(>|t|)"] <= .05
    } else sigs[i] <- cor.test(model.data[, "y"], model.data[, c(...)[i]])$p.value <= .05
  }
  # If a relationship is found, extract the residuals from a regression
  if (any(sigs)) {
    sig_covs <- c(...)[sigs]
    f <- as.formula(paste0("y ~ ", paste(sig_covs, collapse = " + ")))
    # now get the residuals:
    resids <- resid(lm(formula = f, data = model.data[, c("y", sig_covs)]))
  } else resids <- data[, col_name] - mean(data[, col_name]) # just mean-center it instead.
  resids # output the residuals (or the original variable if no relationship was found)
}
```
# In this step, we lose any subjets whose age is not known too
```{r}
data.icvAdj$covAdj <- lapply(data.icvAdj$noCovAdj, function(dat) {
  df <- dat %>% drop_na(age, gender)
  cbind.data.frame(df %>% select(IDENT_SUBID:starts_with("CBCL")),
                   setNames(as.data.frame(
                     lapply(names(df %>% select(-IDENT_SUBID:-starts_with("CBCL"))), function(col) {
                       adjustFun(col_name = col, "age", "gender", data = df)
                      })
                   ), nm = names(df %>% select(-IDENT_SUBID:-starts_with("CBCL")))))
})
```
# Also adjust the structural covariate for age and gender that were not adjusted for ICV
```{r}
data.rawBrain$covAdjusted <- lapply(data.rawBrain$noCovAdj, function(dat) {
  df <- dat %>% drop_na(age, gender)
  cbind.data.frame(df %>% select(IDENT_SUBID:starts_with("CBCL")),
                   setNames(as.data.frame(
                     lapply(names(df %>% select(-IDENT_SUBID:-starts_with("CBCL"))), function(col) {
                       adjustFun(col_name = col, "age", "gender", data = df)
                      })
                   ), nm = names(df %>% select(-IDENT_SUBID:-starts_with("CBCL")))))
})
```
# As a final step, delete any rows for whom the response does not exist
```{r}
# data.icvAdj$covAdj <- lapply(data.icvAdj$covAdj, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
# data.icvAdj$noCovAdj <- lapply(data.icvAdj$noCovAdj, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
# data.rawBrain$noCovAdj <- lapply(data.rawBrain$noCovAdj, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
# data.rawBrain$covAdjusted <- lapply(data.rawBrain$covAdjusted, function(x){
#   x %>% drop_na(starts_with("CBCL"))
# })
```

# Fit models for the ICV-adjusted predictors (with and without adjustement for demographics)
```{r}
# prep the data by dropping NA from the response variable
rfDF.icvAdj <- list()
rfDF.icvAdj$noCovAdj <- lapply(data.icvAdj$noCovAdj, function(d) d %>% drop_na(starts_with("CBCL")))
rfDF.icvAdj$covAdj <- lapply(data.icvAdj$covAdj, function(d) d %>% drop_na(starts_with("CBCL")))
rfDF.rawBrain <- list()
rfDF.rawBrain$noCovAdj <- lapply(data.rawBrain$noCovAdj, function(d) d %>% drop_na(starts_with("CBCL")))
rfDF.rawBrain$covAdjusted <-lapply(data.rawBrain$covAdjusted, 
                                   function(d) d %>% drop_na(starts_with("CBCL")))
# define function to fit the model
# create an option to include or drop the last column (which ICV where it exists)
rf_func <- function(df, icv = FALSE)
{
  model.data <- df %>%
    select(y = starts_with("CBCL"), Left.Thalamus.Proper:last_col(offset = as.numeric(icv)))
  randomForest(y ~ ., data = model.data,
               mtry = 6, ntree = 1000, importance = T, strata = df$GROUP)
}
# Fit the models
icv.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_func)})
raw.withICV.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_func)})
raw.noICV.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, function(x) rf_func(x, icv=TRUE))})
```
# Results
```{r}
rf_model_list <- list(
  "icvAdjusted" = icv.rf_models,
  "notAdjustedByICV" = raw.withICV.rf_models,
  "notICVAdjustedNotInclAsPredictor" = raw.noICV.rf_models
)
saveRDS(rf_model_list, "../../../CBCL_pipeline/output/rf_models_allVars.rds")
lapply(rf_model_list, function(results) {lapply(results, function(list) {lapply(list, function(rf) rf)})})
```

# Re-fit the models on the best variables
```{r}
training.data <- raw_str_csv %>%
  select(SUBJECTID_long, all_of(names(brainData[-1:-3]))) %>%
  filter(!grepl("fu2$", SUBJECTID_long)) %>%
  filter(rowSums(select(., -SUBJECTID_long), na.rm = T)  != 0) %>%
  mutate(IDENT_SUBID = substr(SUBJECTID_long, 1, 5)) %>% select(-SUBJECTID_long) %>%
  filter(!duplicated(IDENT_SUBID)) %>%
  select(IDENT_SUBID, everything()) %>%
  merge(., 
        cbcl %>% 
          select(., IDENT_SUBID, GROUP) %>%
          filter(!duplicated(IDENT_SUBID)) %>%
          filter(!is.na(GROUP)), 
        by = "IDENT_SUBID") %>%
  mutate_at("GROUP", ~factor(., levels = c("PI", "COMP")), all.y = F) %>%
  select(-EstimatedTotalIntraCranialVol)
# fit random forest
rf.train <- randomForest(x = select(training.data, -GROUP, -IDENT_SUBID),
                         y = training.data$GROUP,
                         mtry = 6, ntree = 1000, importance = T, strata = training.data$GROUP,
                         sampsize = rep(sum(training.data$GROUP=="PI"), times = 2))
# permute var imps
pimpRes <- vita::PimpTest(PIMP(X = select(training.data, -GROUP, -IDENT_SUBID), 
                               y = training.data$GROUP,
                               rForest = rf.train,
                               S = 1000))
saveRDS(pimpRes, "../../../CBCL_pipeline/output/pimpResults_structure.rds")
data.frame(pimpRes$pvalue, "var"=rownames(pimpRes$pvalue), stringsAsFactors = FALSE) %>% 
  arrange(p.value) %>% .[1:11,"var"]
```

refit the models with the top 11 of 21 variables
```{r}
bestVars <- data.frame(pimpRes$pvalue, "var"=rownames(pimpRes$pvalue), stringsAsFactors = FALSE) %>% 
  arrange(p.value) %>% .[1:11,"var"]
rf_funcBest <- function(df, icv = FALSE)
{
  model.data <- df %>% select(y = starts_with("CBCL"), all_of(bestVars))
  randomForest(y ~ ., data = model.data,
               mtry = 8, ntree = 1000, importance = T, strata = df$GROUP)
} 
bestVar.icv.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_funcBest)})
bestVar.raw.withICV.rf_models <- lapply(rfDF.icvAdj, function(list) {lapply(list, rf_funcBest)})
```
```{r}
bestVars_rf_model_list <- list(
  "icvAdjusted" = bestVar.icv.rf_models,
  "notAdjustedByICV" = bestVar.raw.withICV.rf_models
)
saveRDS(bestVars_rf_model_list, "../../../CBCL_pipeline/output/bestVars_rf_model_list.rds")
lapply(bestVars_rf_model_list, function(list) {lapply(list, function(rf) rf)})
```

# Variable importance plots
```{r}
source("../../../misc/vimPlot.R")
rf_model_list <- readRDS("../../../CBCL_pipeline/output/rf_models_allVars.rds")
lapply(rf_model_list, function(list) {lapply(list, function(list2) {lapply(list2, vimPlot)})})
```
--->
