---
title: "finalModel_gridCVResults"
author: "Ian Douglas"
date: "5/7/2020"
output: html_document
---
```{r}
library(tidyverse)
```
### Read in the grid results and the data used for each one
```{r}
grid.res <- readRDS("../../output/grid_search/noCC_finalCBCL_ModelsGridCV_2020-05-13.rds")
final.df.list <- readRDS("~/DANL/SB/ianMasters/data/processed/finalStrAdjDF-LIST_2020-05-07.rds")
```
# re create the grid used to generate the results
```{r}
df.list <- lapply(final.df.list, function(df) {
  df %>% 
    rename_at(vars(contains("CBCL")), ~replace(., TRUE, "y")) %>%
    select(-GENDER_FEMALE, -brain_age_yrs, -WAVE, -subjWeight, -EstimatedTotalIntraCranialVol) %>%
    # Now delete all the corpus callosum variables!
    select(-starts_with('CC_')) %>%
    column_to_rownames("IDENT_SUBID")
})
# Now regenerate the grid identically to the way it was during the grid search
# Note that in the cross validation script the data is also contained in the below list
# Here we just need the grid again, not in a nested object with the data too.
grd.list = lapply(df.list, function(df) {
  expand.grid(
      .mtry = 2:(ncol(df) - 3), # minus the response and GROUP; the maximum is p - 1
      .ntree = seq(500, 1550, by = 75),
      .nodesize = seq(1, 20, by = 1)
  )
})
```
# First the Pipeline without Corpus Callosum
```{r,eval=F, echo=F}
all.rsq.list <- lapply(names(grid.res), function(nm) {
  sapply(1:nrow(grd.list[[nm]]), function(i) {
    tail(grid.res[[nm]][[i]]$rsq, 1)
  })
})
names(all.rsq.list) <- names(grid.res)
all.mse.list <- lapply(names(grid.res), function(nm) {
  sapply(1:nrow(grd.list[[nm]]), function(i) {
    tail(grid.res[[nm]][[i]]$mse, 1)
  })
})
names(all.mse.list) <- names(grid.res)
```
Best result based on rsq and mse (seperately)
```{r, eval=F, echo=F}
lapply(all.rsq.list, which.max)
lapply(all.mse.list, which.min)
```
# Manually input the number of the best iteration so as to identify the best hyperparameters
```{r, eval=F}
rbind(
data.frame(response = "internalizing",
           grd.list$CBCL_INTERNALIZING_T[211,],
           rsq = all.rsq.list$CBCL_INTERNALIZING_T[211],
           mse = all.mse.list$CBCL_INTERNALIZING_T[211]),
data.frame(response = "externalizing",
           grd.list$CBCL_EXTERNALIZING_T[3823,],
           rsq = all.rsq.list$CBCL_EXTERNALIZING_T[3823],
           mse = all.mse.list$CBCL_EXTERNALIZING_T[3823]),
data.frame(response = "total",
           grd.list$CBCL_TOTALPROB_T[3165,],
           rsq = all.rsq.list$CBCL_TOTALPROB_T[3165],
           mse = all.mse.list$CBCL_TOTALPROB_T[3165])
) -> fnl.mod.tbl
saveRDS(fnl.mod.tbl, "../../output/noCC_FINAL-bestHyperparameters-2020-05-10.rds")
```
```{r}
fnl.mod.tbl <- readRDS("../../output/noCC_FINAL-bestHyperparameters-2020-05-10.rds")
fnl.mod.tbl
```

Iteratively create models to reproduce the above
Again, manually input the best parameters from the above:
```{r, eval=F}
# define functions to automatically produce these forests with different seeds
RF_intrnl <- function(seed)
{
  set.seed(seed)
  randomForest(
    y ~ .-GROUP,
    data = df.list$CBCL_INTERNALIZING_T,
    strata = df.list$CBCL_INTERNALIZING_T$GROUP,
    sampsize = sum(
      df.list$CBCL_INTERNALIZING_T$GROUP=="PI"
      )*n_distinct(df.list$CBCL_INTERNALIZING_T$GROUP),
  # save desired outputs:
  importance = T, proximity = T, keep.forest = T, localImp = T,
  # optimal parameters:
  mtry = 2, ntree = 500, nodesize = 2
  )
}

RF_extrnl <- function(seed) 
{
  set.seed(seed)
  randomForest(
    y ~ .-GROUP,
    data = df.list$CBCL_EXTERNALIZING_T,
    strata = df.list$CBCL_EXTERNALIZING_T$GROUP,
    sampsize = sum(
      df.list$CBCL_EXTERNALIZING_T$GROUP=="PI"
      )*n_distinct(df.list$CBCL_EXTERNALIZING_T$GROUP),
    # save desired outputs:
    importance = T, proximity = T, keep.forest = T, localImp = T,
    # optimal parameters:
    mtry = 2, ntree = 725, nodesize = 19
  )
}

RF_ttl <- function(seed, reset.seed = TRUE)
{
  if (reset.seed) set.seed(seed)
  randomForest(
    y ~ .-GROUP,
    data = df.list$CBCL_TOTALPROB_T,
    strata = df.list$CBCL_TOTALPROB_T$GROUP,
    sampsize = sum(
      df.list$CBCL_TOTALPROB_T$GROUP=="PI"
      )*n_distinct(df.list$CBCL_TOTALPROB_T$GROUP),
    # save desired outputs:
    importance = T, proximity = T, keep.forest = T, localImp = T, keep.inbag = T,
    # optimal parameters:
    mtry = 2, ntree = 575, nodesize = 16
  )
}
```

# create forests with these parameters with different seeds until the optimal model is found
Here, manually input the RSQ we are searching for
```{r, eval = F}
# First for internalizing cbcl
rf.intrnl <- NULL
seed <- 0
RSQ <- -Inf
t0 <- Sys.time()
while (RSQ < 0.0530) {
  seed <- seed + 1
  rf.intrnl <- RF_intrnl(seed)
  RSQ <- tail(rf.intrnl$rsq, 1)
  t1 <- Sys.time()
}
print("Internalizing:")
print(paste0("Elapsed Time: ", t1 - t0))
print(paste0("Best model found at seed: ", seed))

# Externalizing cbcl
rf.extrnl <- NULL
seed <- 0
RSQ <- -Inf
t0 <- Sys.time()
while (RSQ < -0.030) {
  seed <- seed + 1
  rf.extrnl <- RF_extrnl(seed)
  RSQ <- tail(rf.extrnl$rsq, 1)
  t1 <- Sys.time()
}
print("Externalizing:")
print(paste0("Elapsed Time: ", t1 - t0))
print(paste0("Best model found at seed: ", seed))

# Total cbcl
# This one never reached its "optimal" value above, after hundreds of thousands of tries, so I am just going to take the model with the same same hyperparameters out of 100 tries with the best RSQ
# rf.ttl <- NULL
# seed <- 0
# RSQ <- -Inf
# t0 <- Sys.time()
# while (RSQ < -0.01) {
#   seed <- seed + 1
#   rf.ttl <- RF_ttl(seed)
#   RSQ <- tail(rf.ttl$rsq, 1)
#   t1 <- Sys.time()
# }
# print("Total")
# print(paste0("Elapsed Time: ", t1 - t0))
# print(paste0("Best model found at seed: ", seed))
rf.ttl <- NULL
rfi <- NULL
RSQ <- NULL
RSQi <- NULL
set.seed(111)
for (i in 1:100) {
  if (i == 1) {
    rf.ttl <- RF_ttl(reset.seed = FALSE)
    RSQ <- tail(rf.ttl$rsq, 1)
  } else {
    rfi <- RF_ttl(reset.seed = FALSE)
    RSQi <- tail(rfi$rsq, 1)
    if (RSQi > RSQ) {
      rf.ttl <- rfi
      RSQ <- RSQi
    }
  }
}
```
# Save final models
```{r, eval=F}
saveRDS(rf.intrnl, "../../output/cbcl_pipeline/noCC_FINALMODEL-cbclInternalizing-2020-05-10.rds")
saveRDS(rf.extrnl, "../../output/cbcl_pipeline/noCC_FINALMODEL-cbclExternalizing-2020-05-10.rds")
saveRDS(rf.ttl, "../../output/cbcl_pipeline/noCC_FINALMODEL-cbclTotalprob-2020-05-10.rds")
```
```{r,eval=T, echo=F}
rf.intrnl <- readRDS("../../output/cbcl_pipeline/noCC_FINALMODEL-cbclInternalizing-2020-05-10.rds")
rf.extrnl <- readRDS("../../output/cbcl_pipeline/noCC_FINALMODEL-cbclExternalizing-2020-05-10.rds")
rf.ttl <- readRDS("../../output/cbcl_pipeline/noCC_FINALMODEL-cbclTotalprob-2020-05-10.rds")
```

# local importances
```{r, eval=F}
localImps <- list(
  'intrnl' = list(
    'PI' = rf.intrnl$localImportance[,final.df.list$CBCL_INTERNALIZING_T$GROUP=="PI"],
    'COMP' = rf.intrnl$localImportance[,final.df.list$CBCL_INTERNALIZING_T$GROUP!="PI"]),
  'extrnl' = list(
    'PI'=rf.extrnl$localImportance[,final.df.list$CBCL_EXTERNALIZING_T$GROUP=="PI"],
    'COMP' = rf.extrnl$localImportance[,final.df.list$CBCL_EXTERNALIZING_T$GROUP!="PI"]),
  'total'= list(
    'PI'= rf.ttl$localImportance[,final.df.list$CBCL_TOTALPROB_T$GROUP == "PI"],
    'COMP' = rf.ttl$localImportance[,final.df.list$CBCL_TOTALPROB_T$GROUP!="PI"])
)
ageSplitLocalImp <- list(
  'intrnl' = list(
    'under11yrsOld' = rf.intrnl$localImportance[,final.df.list$CBCL_INTERNALIZING_T$brain_age_yrs < 11.0],
    'elevenYrsOrOlder' = rf.intrnl$localImportance[,final.df.list$CBCL_INTERNALIZING_T$brain_age_yrs >= 11.0]),
  'extrnl' = list(
    'under11yrsOld'=rf.extrnl$localImportance[,final.df.list$CBCL_EXTERNALIZING_T$brain_age_yrs < 11.0],
    'elevenYrsOrOlder' = rf.extrnl$localImportance[,final.df.list$CBCL_EXTERNALIZING_T$brain_age_yrs >= 11.0]),
  'total'= list(
    'under11yrsOld'= rf.ttl$localImportance[,final.df.list$CBCL_TOTALPROB_T$brain_age_yrs < 11.0],
    'elevenYrsOrOlder' = rf.ttl$localImportance[,final.df.list$CBCL_TOTALPROB_T$brain_age_yrs < 11.0])
)
saveRDS(ageSplitLocalImp, "../../output/cbcl_pipeline/noCC_ageSplitlocalImportancesFINAL-2020-05-10.rds")
saveRDS(localImps,"../../output/cbcl_pipeline/noCC_localImportancesFINAL-2020-05-10.rds")
```

# Variable importance plots
```{r, eval=F}
plt_data <- map_dfr(names(localImps), function(nm) {
  map_dfr(c("PI","COMP"), function(grp) {
    data.frame(
      response = nm,
      group = grp,
      variable = rownames(localImps[[nm]][[grp]]),
      importance = rowMeans(localImps[[nm]][[grp]]),
      stringsAsFactors = F
    )
  })
}) %>%
  mutate_at(vars(response), ~case_when(
    . == "intrnl" ~ 'internalizing', . == "extrnl" ~ "externalizing", TRUE ~ "total"
  )) %>%
  group_by(response, group) %>%
  arrange(variable) %>% 
  mutate_at(vars(variable), ~factor(., levels = rev(unique(.))))
saveRDS(plt_data,
        "../../output/cbcl_pipeline/noCC_plotDataFrom-FINALMODELS-forFINALvarImpPlot-2020-05-10.rds")
```
```{r, echo=F, eval=T}
plt_data <- readRDS(
  "../../output/cbcl_pipeline/noCC_plotDataFrom-FINALMODELS-forFINALvarImpPlot-2020-05-10.rds")
```
### For age
```{r}
age_plt_data <- map_dfr(names(ageSplitLocalImp), function(nm) {
  map_dfr(c("under11yrsOld","elevenYrsOrOlder"), function(grp) {
    data.frame(
      response = nm,
      group = grp,
      variable = rownames(ageSplitLocalImp[[nm]][[grp]]),
      importance = rowMeans(ageSplitLocalImp[[nm]][[grp]]),
      stringsAsFactors = F
    )
  })
}) %>%
  mutate_at(vars(response), ~case_when(
    . == "intrnl" ~ 'internalizing', . == "extrnl" ~ "externalizing", TRUE ~ "total"
  )) %>%
  group_by(response, group) %>%
  arrange(variable) %>% 
  mutate_at(vars(variable), ~factor(., levels = rev(unique(.))))
saveRDS(age_plt_data,
        "../../output/cbcl_pipeline/noCC_ageSplit_plotDataFrom-FINALMODELS-forFINALvarImpPlot-2020-05-10.rds")
```

```{r, fig.asp=.7, fig.width=14}
ggplot(plt_data, aes(x = importance, y = variable)) +
  geom_bar(stat = "identity", fill = "darkblue") + 
  facet_grid(group~response, scales = 'free') +
  theme_linedraw() +
  theme(text = element_text(size = 15),
        plot.title.position = "plot",
        plot.title = element_text(hjust = .5, size = 18),
        strip.text = element_text(size = 16.5)) +
  labs(title = "Group-averaged Local Variable Importances from Random Forest Regressions\n of CBCL Externalizing, Internalizing, and Total Subscales") +
  ylab("Subcortical brain region") +
  xlab("% Change in Mean Squared Error")
# ggsave(filename = "~/DANL/SB/ianMasters/results/RF/plots/FINALMODEL-varimp-CBCLonSTR.jpg",
#        units = "in", width = 14, height = .7*14)
# ggsave(filename = "~/DANL/SB/ianMasters/FINALMODEL-varimp-CBCLonSTR.jpg",
#        units = "in", width = 14, height = .7*14)
```

##For age
```{r}
ggplot(age_plt_data, aes(x = importance, y = variable)) +
  geom_bar(stat = "identity", fill = "darkblue") + 
  facet_grid(group~response, scales = 'free') +
  theme_linedraw() +
  theme(text = element_text(size = 15),
        plot.title.position = "panel",
        plot.title = element_text(hjust = .5, size = 18),
        strip.text = element_text(size = 16.5)) +
  labs(title = "Age-group averaged Local Variable Importances from Random Forest Regressions\n of CBCL Externalizing, Internalizing, and Total Subscales") +
  ylab("Subcortical brain region") +
  xlab("% Change in Mean Squared Error")
ggsave(filename = "~/DANL/SB/ianMasters/results/RF/plots/noCC_ageStrat-FINALMODEL-varimp-CBCLonSTR.jpg",
        units = "in", width = 14, height = .7*14)
ggsave(filename = "~/DANL/SB/ianMasters/noCC_ageStrat-FINALMODEL-varimp-CBCLonSTR.jpg",
        units = "in", width = 14, height = .7*14)
```

```{r}
par(mfrow = c(1, 3))
varImpPlot(rf.intrnl, type = 1, main = "Full sample variable importance\n predicting Internalizing")
varImpPlot(rf.extrnl, type = 1, main = "Full sample variable importance\n predicting Externalizing")
varImpPlot(rf.ttl, type = 1, main = "Full sample variable importance\n predicting Total CBCL")
```


# The pipeline with CC regions
### Read in the grid results and the data used for each one
```{r}
# grid.res <- readRDS("../../output/grid_search/finalCBCL_ModelsGridCV_2020-05-07.rds")
final.df.list <- readRDS("~/DANL/SB/ianMasters/data/processed/finalStrAdjDF-LIST_2020-05-07.rds")
```
# re create the grid used to generate the results
```{r}
df.list <- lapply(final.df.list, function(df) {
  df %>% 
    rename_at(vars(contains("CBCL")), ~replace(., TRUE, "y")) %>%
    select(-GENDER_FEMALE, -WAVE, -brain_age_yrs, -subjWeight, -EstimatedTotalIntraCranialVol) %>%
    column_to_rownames("IDENT_SUBID")
})
# Now regenerate the grid identically to the way it was during the grid search
grd.list <- lapply(df.list, function(df) {
  expand.grid(
    .mtry = 2:(ncol(df) - 3), # minus the response and GROUP; the maximum is p - 1
    .ntree = seq(500, 1550, by = 75),
    .nodesize = seq(1, 20, by = 1)) 
})
```

```{r,eval=F, echo=F}
all.rsq.list <- lapply(names(grid.res), function(nm) {
  sapply(1:nrow(grd.list[[nm]]), function(i) {
    tail(grid.res[[nm]][[i]]$rsq, 1)
  })
})
names(all.rsq.list) <- names(grid.res)
all.mse.list <- lapply(names(grid.res), function(nm) {
  sapply(1:nrow(grd.list[[nm]]), function(i) {
    tail(grid.res[[nm]][[i]]$mse, 1)
  })
})
names(all.mse.list) <- names(grid.res)
```
Best result based on rsq and mse (seperately)
```{r, eval=F, echo=F}
lapply(all.rsq.list, which.max)
lapply(all.mse.list, which.min)
```
```{r, eval=F}
rbind(
data.frame(response = "internalizing",
           grd.list$CBCL_INTERNALIZING_T[1143,],
           rsq = all.rsq.list$CBCL_INTERNALIZING_T[1143],
           mse = all.mse.list$CBCL_INTERNALIZING_T[1143]),
data.frame(response = "externalizing",
           grd.list$CBCL_EXTERNALIZING_T[4846,],
           rsq = all.rsq.list$CBCL_EXTERNALIZING_T[4846],
           mse = all.mse.list$CBCL_EXTERNALIZING_T[4846]),
data.frame(response = "total",
           grd.list$CBCL_TOTALPROB_T[2935,],
           rsq = all.rsq.list$CBCL_TOTALPROB_T[2935],
           mse = all.mse.list$CBCL_TOTALPROB_T[2935])
) -> fnl.mod.tbl
# saveRDS(fnl.mod.tbl, "../../output/FINAL-bestHyperparameters-2020-05-10.rds")
```
```{r}
fnl.mod.tbl <- readRDS("../../output/FINAL-bestHyperparameters-2020-05-10.rds")
fnl.mod.tbl
```

Iteratively create models to reproduce the above
```{r, eval=F}
# define functions to automatically produce these forests with different seeds
RF_intrnl <- function(seed)
{
  set.seed(seed)
  randomForest(
    y ~ .-GROUP,
    data = df.list$CBCL_INTERNALIZING_T,
    strata = df.list$CBCL_INTERNALIZING_T$GROUP,
    sampsize = sum(
      df.list$CBCL_INTERNALIZING_T$GROUP=="PI"
      )*n_distinct(df.list$CBCL_INTERNALIZING_T$GROUP),
  # save desired outputs:
  importance = T, proximity = T, keep.forest = T, localImp = T,
  # optimal parameters:
  mtry = 4, ntree = 500, nodesize = 5
  )
}

RF_extrnl <- function(seed) 
{
  set.seed(seed)
  randomForest(
    y ~ .-GROUP,
    data = df.list$CBCL_EXTERNALIZING_T,
    strata = df.list$CBCL_EXTERNALIZING_T$GROUP,
    sampsize = sum(
      df.list$CBCL_EXTERNALIZING_T$GROUP=="PI"
      )*n_distinct(df.list$CBCL_EXTERNALIZING_T$GROUP),
    # save desired outputs:
    importance = T, proximity = T, keep.forest = T, localImp = T,
    # optimal parameters:
    mtry = 2, ntree = 500, nodesize = 18
  )
}

RF_ttl <- function(seed)
{
  set.seed(seed)
  randomForest(
    y ~ .-GROUP,
    data = df.list$CBCL_TOTALPROB_T,
    strata = df.list$CBCL_TOTALPROB_T$GROUP,
    sampsize = sum(
      df.list$CBCL_TOTALPROB_T$GROUP=="PI"
      )*n_distinct(df.list$CBCL_TOTALPROB_T$GROUP),
    # save desired outputs:
    importance = T, proximity = T, keep.forest = T, localImp = T, keep.inbag = T,
    # optimal parameters:
    mtry = 10, ntree = 800, nodesize = 11
  )
}
```

# create forests with these parameters with different seeds until the optimal model is found
```{r, eval = F}
# First for internalizing cbcl
rf.intrnl <- NULL
seed <- 1
RSQ <- -Inf
t0 <- Sys.time()
while (RSQ < .1340) {
  rf.intrnl <- RF_intrnl(seed)
  RSQ <- tail(rf.intrnl$rsq, 1)
  seed <- seed + 1
  t1 <- Sys.time()
}
print(paste0("Elapsed Time: ", t1 - t0))
print(paste0("Best model found at seed: ", seed))
```
11.5 seconds, seed is 54

```{r, eval = F}
# Externalizing cbcl
rf.extrnl <- NULL
seed <- 1
RSQ <- -Inf
t0 <- Sys.time()
while (RSQ < .0260) {
  rf.extrnl <- RF_extrnl(seed)
  RSQ <- tail(rf.extrnl$rsq, 1)
  seed <- seed + 1
  t1 <- Sys.time()
}
print(paste0("Elapsed Time: ", t1 - t0))
print(paste0("Best model found at seed: ", seed))
```
1.38 minutes, seed is 709.

```{r, eval = F}
# Total cbcl
rf.ttl <- NULL
seed <- 1
RSQ <- -Inf
t0 <- Sys.time()
while (RSQ < .0869) {
  rf.ttl <- RF_ttl(seed)
  RSQ <- tail(rf.ttl$rsq, 1)
  seed <- seed + 1
  t1 <- Sys.time()
}
print(paste0("Elapsed Time: ", t1 - t0))
print(paste0("Best model found at seed: ", seed - 1)) # the printout below is wrong!
```
12.29 minutes, seed is 1520

# Save final models
```{r, eval=F}
saveRDS(rf.intrnl, "../../output/cbcl_pipeline/FINALMODEL-cbclInternalizing-2020-05-10.rds")
saveRDS(rf.extrnl, "../../output/cbcl_pipeline/FINALMODEL-cbclExternalizing-2020-05-10.rds")
saveRDS(rf.ttl, "../../output/cbcl_pipeline/FINALMODEL-cbclTotalprob-2020-05-10.rds")
```
```{r,eval=T, echo=F}
rf.intrnl <- readRDS("../../output/cbcl_pipeline/FINALMODEL-cbclInternalizing-2020-05-10.rds")
rf.extrnl <- readRDS("../../output/cbcl_pipeline/FINALMODEL-cbclExternalizing-2020-05-10.rds")
rf.ttl <- readRDS("../../output/cbcl_pipeline/FINALMODEL-cbclTotalprob-2020-05-10.rds")
```

# local importances
```{r, eval=F}
localImps <- list(
  'intrnl' = list(
    'PI' = rf.intrnl$localImportance[,df.list$CBCL_INTERNALIZING_T$GROUP=="PI"],
    'COMP' = rf.intrnl$localImportance[,df.list$CBCL_INTERNALIZING_T$GROUP!="PI"]),
  'extrnl' = list(
    'PI'=rf.extrnl$localImportance[,df.list$CBCL_EXTERNALIZING_T$GROUP=="PI"],
    'COMP' = rf.extrnl$localImportance[,df.list$CBCL_EXTERNALIZING_T$GROUP!="PI"]),
  'total'= list(
    'PI'= rf.ttl$localImportance[,df.list$CBCL_TOTALPROB_T$GROUP == "PI"],
    'COMP' = rf.ttl$localImportance[,df.list$CBCL_TOTALPROB_T$GROUP!="PI"])
)
ageSplitLocalImp <- list(
  'intrnl' = list(
    'youngChild' = rf.intrnl$localImportance[,df.list$CBCL_INTERNALIZING_T$brain_age_yrs < 11.0],
    'adolescent' = rf.intrnl$localImportance[,df.list$CBCL_INTERNALIZING_T$brain_age_yrs >= 11.0]),
  'extrnl' = list(
    'youngChild'=rf.extrnl$localImportance[,df.list$CBCL_EXTERNALIZING_T$brain_age_yrs < 11.0],
    'adolescent' = rf.extrnl$localImportance[,df.list$CBCL_EXTERNALIZING_T$brain_age_yrs >= 11.0]),
  'total'= list(
    'youngChild'= rf.ttl$localImportance[,df.list$CBCL_TOTALPROB_T$brain_age_yrs < 11.0],
    'adolescent' = rf.ttl$localImportance[,df.list$CBCL_TOTALPROB_T$brain_age_yrs < 11.0])
)
saveRDS(ageSplitLocalImp, "../../output/cbcl_pipeline/ageSplitlocalImportancesFINAL-2020-05-10.rds")
saveRDS(localImps,"../../output/cbcl_pipeline/localImportancesFINAL-2020-05-10.rds")
```

# Variable importance plots
```{r, eval=F}
plt_data <- map_dfr(names(localImps), function(nm) {
  map_dfr(c("PI","COMP"), function(grp) {
    data.frame(
      response = nm,
      group = grp,
      variable = rownames(localImps[[nm]][[grp]]),
      importance = rowMeans(localImps[[nm]][[grp]]),
      stringsAsFactors = F
    )
  })
}) %>%
  mutate_at(vars(response), ~case_when(
    . == "intrnl" ~ 'internalizing', . == "extrnl" ~ "externalizing", TRUE ~ "total"
  )) %>%
  group_by(response, group) %>%
  arrange(variable) %>% 
  mutate_at(vars(variable), ~factor(., levels = rev(unique(.))))
saveRDS(plt_data,
        "../../output/cbcl_pipeline/plotDataFrom-FINALMODELS-forFINALvarImpPlot-2020-05-10.rds")
```
```{r, echo=F, eval=T}
plt_data <- readRDS(
  "../../output/cbcl_pipeline/plotDataFrom-FINALMODELS-forFINALvarImpPlot-2020-05-10.rds")
```
### For age
```{r}
age_plt_data <- map_dfr(names(ageSplitLocalImp), function(nm) {
  map_dfr(c("youngChild","adolescent"), function(grp) {
    data.frame(
      response = nm,
      group = grp,
      variable = rownames(ageSplitLocalImp[[nm]][[grp]]),
      importance = rowMeans(ageSplitLocalImp[[nm]][[grp]]),
      stringsAsFactors = F
    )
  })
}) %>%
  mutate_at(vars(response), ~case_when(
    . == "intrnl" ~ 'internalizing', . == "extrnl" ~ "externalizing", TRUE ~ "total"
  )) %>%
  group_by(response, group) %>%
  arrange(variable) %>% 
  mutate_at(vars(variable), ~factor(., levels = rev(unique(.))))
saveRDS(age_plt_data,
        "../../output/cbcl_pipeline/ageSplit_plotDataFrom-FINALMODELS-forFINALvarImpPlot-2020-05-10.rds")
```

```{r, fig.asp=.7, fig.width=14}
ggplot(plt_data, aes(x = importance, y = variable)) +
  geom_bar(stat = "identity", fill = "darkblue") + 
  facet_grid(group~response, scales = 'free') +
  theme_linedraw() +
  theme(text = element_text(size = 15),
        plot.title.position = "plot",
        plot.title = element_text(hjust = .5, size = 18),
        strip.text = element_text(size = 16.5)) +
  labs(title = "Group-averaged Local Variable Importances from Random Forest Regressions\n of CBCL Externalizing, Internalizing, and Total Subscales") +
  ylab("Subcortical brain region") +
  xlab("% Change in Mean Squared Error")
# ggsave(filename = "~/DANL/SB/ianMasters/results/RF/plots/FINALMODEL-varimp-CBCLonSTR.jpg",
#        units = "in", width = 14, height = .7*14)
# ggsave(filename = "~/DANL/SB/ianMasters/FINALMODEL-varimp-CBCLonSTR.jpg",
#        units = "in", width = 14, height = .7*14)
```

##For age
```{r}
ggplot(age_plt_data, aes(x = importance, y = variable)) +
  geom_bar(stat = "identity", fill = "darkblue") + 
  facet_grid(group~response, scales = 'free') +
  theme_linedraw() +
  theme(text = element_text(size = 15),
        plot.title.position = "plot",
        plot.title = element_text(hjust = .5, size = 18),
        strip.text = element_text(size = 16.5)) +
  labs(title = "Age-group averaged Local Variable Importances from Random Forest Regressions\n of CBCL Externalizing, Internalizing, and Total Subscales") +
  ylab("Subcortical brain region") +
  xlab("% Change in Mean Squared Error")
ggsave(filename = "~/DANL/SB/ianMasters/results/RF/plots/ageStrat-FINALMODEL-varimp-CBCLonSTR.jpg",
        units = "in", width = 14, height = .7*14)
```

```{r}
par(mfrow = c(1, 3))
varImpPlot(rf.intrnl, type = 1, main = "Full sample variable importance\n predicting Internalizing")
varImpPlot(rf.extrnl, type = 1, main = "Full sample variable importance\n predicting Externalizing")
varImpPlot(rf.ttl, type = 1, main = "Full sample variable importance\n predicting Total CBCL")
```

