---
title: "structuralData_adjustmentForDemographics"
author: "Ian Douglas"
date: "4/7/2020"
output: html_document
---
### Load in the packages required
```{r}
library(tidyverse)
```

### Read in the structural volume MRI
```{r}
raw_str_csv <- read.csv("../../data/raw/structural/v6_asegstats_all_compiled_Marta_02-01-2018.csv")

#### Reformatting
rawStructural <- raw_str_csv %>%
  # Before using regular expression functions, manage the case sensitivity
  mutate_at(vars(SUBJECTID_long), ~sub("sb", "SB", tolower(.))) %>%
  # Firstly,delete any structural data that was collected at wave three as a different scanner was used
  filter(!grepl("fu2", SUBJECTID_long)) %>%
  # create variables that exist in the master which indcating which wave of data to pull
  mutate(IDENT_SUBID = sub("_fu[0-9]", "", SUBJECTID_long),
         index_wave = ifelse(grepl("_fu", SUBJECTID_long), as.numeric(substr(SUBJECTID_long, 9, 9)) + 1, 1),
         uniq_key = paste0(IDENT_SUBID, "_", index_wave)) %>%
  ungroup() %>%
  select(IDENT_SUBID, uniq_key, everything()) %>%
  # also deselect several variables that were used to compute other volumes; redundant ID and index wave columns
  select(-Left.WM.hypointensities:-Right.non.WM.hypointensities, -X5th.Ventricle, -SUBJECTID_long, -index_wave)
```
### Read in the labels indicating which wave of data to pull
```{r}
labels <- readRDS("../../data/labels/wave1Labels.rds")
```

### Read in the master data frame with the gender variable
```{r}
load( # loads a data frame called "monster_SB"
  "../../data/raw/3_ALL_long_cort_symptoms_brain_structure_function_epi_ages4-19_2019-05-15.Rdata"
)
```

# Merging
```{r}
df1 <- monster_SB %>%
  select(IDENT_SUBID, index_wave, age, gender, GROUP) %>%
  mutate(uniq_key = paste0(IDENT_SUBID, "_", index_wave)) %>%
  select(-IDENT_SUBID, -index_wave) %>%
  merge(x = labels %>% select(IDENT_SUBID, uniq_key = uniq_key_str), y = ., by = "uniq_key", all.x = T) %>%
  merge(x = ., y = rawStructural, by = c("uniq_key", "IDENT_SUBID"), all.x = T)

### Delete those individuals who do not have complete structural MRI data
df2 <- df1 %>% 
  drop_na(Left.Lateral.Ventricle:CortexVol) %>%
  # recode the gender variable
  mutate_at(vars(gender), ~as.factor(ifelse(gender == "Female", 1, 0))) %>%
  rename(GENDER_FEMALE = gender)
saveRDS(df2,"../../data/processed/withNA_StrLabelled.rds")
```

# Combining regions
According to Michelle Van Tiegham's notes,  delete several variables that pertain to overlapping brain regions
```{r}
regions_key <- read.csv("../../data/structuralVarNames20191002_MVT.csv") %>%
  rename_at(2, ~replace(., TRUE, "keep")) %>%
  mutate_at("keep", ~ifelse(. == "X", FALSE, TRUE)) %>%
  filter(keep)

df3 <- df2 %>%
  select(uniq_key:GROUP, all_of(regions_key$var),
         # additionally, drop the whole brain volum measure since it is redundant
         -EstimatedTotalIntraCranialVol)
saveRDS(df3,"../../data/processed/withNA_StrLabelled_SegmentationSubset.rds")
```

<!-- # Clean up some loose ends -->
<!-- Some subjects have wave 2 structural scans, but are mislabeled in the master, or just have index_wave==1 and one row of data. As a result, during merging the information that comes from the master (group, gender, age) did not merge. So fix these here: -->
<!-- ```{r} -->
<!-- ### Finally, delete those who have NA for gender and age, since that is required to adjust the data -->
<!-- ### Additionally if they don't have the GROUP variable, they should be dropped as well. -->

<!-- # This is true of SB150, SB308, and SB312 -->
<!-- df3 <- df2 -->

<!-- ``` -->

# Adjusting by gender and age
```{r}
adjustedStr <- readRDS("../../data/processed/withNA_StrLabelled.rds") %>%
  drop_na()
adjustFun <- function(col_name)
{
  model.data <- setNames(adjustedStr[c(col_name, "age", "GENDER_FEMALE")],
                         nm = c("y", "age", "GENDER_FEMALE")) # rename the covariate to "y"
  # Test for a sig covariance with age and gender:
  corSig = cor.test(model.data$y, model.data$age)$p.value <= .05
  anovaSig = summary(lm(y ~ GENDER_FEMALE, data = model.data))$coefficients[2, "Pr(>|t|)"] <= .05
  # If a relationship is found, extract the residuals from a regression
  if (any(c(corSig, anovaSig))) {
    # record if the variable covaries with age and/or gender
    covariances <- get("covariances", envir = .GlobalEnv)
    covariances[covariances$var==col_name, c("age_cov","gender_cov")] <- c(corSig, anovaSig)
    assign("covariances", value = covariances, pos = .GlobalEnv)
    # now get the residuals:
    resids <- resid(lm(y ~ ., data = model.data[, c("y", c("age","GENDER_FEMALE")[c(corSig, anovaSig)])]))
  } else resids <- adjustedStr[, col_name]
  resids # output the residuals (or the original variable if no relationship was found)
}
## create the 
covariances <- data.frame("var" = names(adjustedStr[-1:-5]), 
                          matrix(FALSE, nrow = length(names(adjustedStr[-1:-5])), ncol= 2,
                                 dimnames = list(NULL, c("age_cov","gender_cov"))))
# Run the function on each column
adjustedStr[-1:-5] <- lapply(names(adjustedStr[-1:-5]), adjustFun)
saveRDS(adjustedStr, "../../data/processed/adjustedStructuralData_2020-04-08.rds")
# process results 
covariances <- covariances %>% mutate(covariate = case_when(
  age_cov == FALSE & gender_cov == FALSE ~ "None",
  age_cov == TRUE & gender_cov == FALSE ~ "Age",
  age_cov == FALSE & gender_cov == TRUE ~ "Gender",
  age_cov == TRUE & gender_cov == TRUE ~ "Both"
))
write.csv(covariances, "../../data/ageGenderCovariationTable_structuralPredictors.csv", row.names = FALSE)
```

# Repeat the adjustment process with subsetted data
```{r}
# read in data and rmove na
adjustedStrSubset <- readRDS(
  "../../data/processed/withNA_StrLabelled_SegmentationSubset.rds") %>%
  drop_na()
# define adjustment function
adjustFun <- function(col_name)
{
  model.data <- setNames(adjustedStrSubset[c(col_name, "age", "GENDER_FEMALE")],
                         nm = c("y", "age", "GENDER_FEMALE")) # rename the covariate to "y"
  # Test for a sig covariance with age and gender:
  corSig = cor.test(model.data$y, model.data$age)$p.value <= .05
  anovaSig = summary(lm(y ~ GENDER_FEMALE, data = model.data))$coefficients[2, "Pr(>|t|)"] <= .05
  # If a relationship is found, extract the residuals from a regression
  if (any(c(corSig, anovaSig))) {
    # record if the variable covaries with age and/or gender
    covariances <- get("covariances", envir = .GlobalEnv)
    covariances[covariances$var==col_name, c("age_cov","gender_cov")] <- c(corSig, anovaSig)
    assign("covariances", value = covariances, pos = .GlobalEnv)
    # now get the residuals:
    resids <- resid(lm(y ~ ., data = model.data[, c("y", c("age","GENDER_FEMALE")[c(corSig, anovaSig)])]))
  } else resids <- adjustedStrSubset[, col_name]
  resids # output the residuals (or the original variable if no relationship was found)
}
## create the 
covariances <- data.frame("var" = names(adjustedStrSubset[-1:-5]), 
                          matrix(FALSE, nrow = length(names(adjustedStrSubset[-1:-5])), ncol= 2,
                                 dimnames = list(NULL, c("age_cov","gender_cov"))))
# Run the function on each column
adjustedStrSubset[-1:-5] <- lapply(names(adjustedStrSubset[-1:-5]), adjustFun)
saveRDS(adjustedStrSubset, "../../data/processed/adjustedStructural_segmentationSubset_2020-04-08.rds")
# process results 
covariances <- covariances %>% mutate(covariate = case_when(
  age_cov == FALSE & gender_cov == FALSE ~ "None",
  age_cov == TRUE & gender_cov == FALSE ~ "Age",
  age_cov == FALSE & gender_cov == TRUE ~ "Gender",
  age_cov == TRUE & gender_cov == TRUE ~ "Both"
))
write.csv(covariances, "../../data/ageGenderCovariationTable_segmentationSubsetStructuralPredictors.csv", row.names = FALSE)
```

