---
title: "PCA"
author: "Ian Douglas"
date: "9/25/2019"
output:
  html_document:
    number_sections: yes
    toc: yes
    df_print: paged
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
    df_print: paged
---
```{r}
library(corrplot)
library(tidyverse)
```

# Load in the functional connectivity and dissimilarity data
-Also read in the labels for plotting purposes
```{r}
fcon = readRDS('~/DANL/SB/data/processed/labelledFCor.rds')
fdis = readRDS('~/DANL/SB/data/processed/labelledFDiss.rds')
fdisBeta1 = readRDS('~/DANL/SB/data/processed/labelledFDissBeta1.rds')
if (nrow(fcon) == nrow(fdis) & ncol(fcon) == ncol(fdis)) 
  {dim(list(fcon, fdis)[[sample(1:2,1)]])}
```

# Conduct principal components analysis
-Note that there are more variables than participants, so the lowest dimensional space in which to project the principal components is equal to $n$, not $p$.
```{r}
fcPCA = prcomp(select(fcon, -age, -GROUP, -IDENT_SUBID), 
                 scale. = TRUE, center = TRUE, retx = TRUE)
saveRDS(funcPCA, '../results/PCA/fcPCAobj.rds')
fdPCA = prcomp(select(fdis, -age, -GROUP, -IDENT_SUBID), 
                 scale. = TRUE, center = TRUE, retx = TRUE)
saveRDS(distPCA, '../results/PCA/fdPCAobj.rds')
fdPCAbeta1 = prcomp(select(fdisBeta1, -age, -GROUP, -IDENT_SUBID),
                    scale. = TRUE, center = TRUE, retx = TRUE)
saveRDS(distPCA, '../results/PCA/fdPCAobj.rds')
# Write out the labelled PC scores data frame
fc.labelledPCA = cbind(select(fcon, IDENT_SUBID,GROUP,age), fcPCA$x)
fd.labelledPCA = cbind(select(fdis, IDENT_SUBID,GROUP,age), fdPCA$x)
fdBeta1.labelledPCA = cbind(select(fdisBeta1, IDENT_SUBID,GROUP,age), fdPCAbeta1$x)
saveRDS(fc.labelledPCA,'~/DANL/SB/data/processed/fcPCASCoresLbl.rds')
saveRDS(fd.labelledPCA,'~/DANL/SB/data/processed/fdPCASCoresLbl.rds')
saveRDS(fdBeta1.labelledPCA,'~/DANL/SB/data/processed/fdBeta1PCASCoresLbl.rds')
```

Explanation: the `fcPCA` is produced from a set of variables, for which each participant has one score for each pair of brain regions (combinations of size 2 of 69 different units) summing to 2346 variables. Each score is the Pearson correlation between the preprocessed time series of each brain region with that of each other brain region.

The `fdPCA` object is derived from the same process, except that instead of computing the correlation between each time series, a measure of the dissimilarity between the two time series is calculated which is based on pearson correlation. For this reason the results are expected to be similar.

### Correlation between the solutions
```{r}
pcCorr = cor(cbind(fcPCA$x, fdPCA$x))
nm = c(paste0("fcPC",1:ncol(fcPCA$x)),
       paste0("fdPC",1:ncol(fdPCA$x)))
rownames(pcCorr) <- nm; colnames(pcCorr) <- nm
methodCompareCor <- as.matrix(pcCorr[1:149, 150:298])
# visualize the correlation between the first 30 dimensions across each method 
corrplot(methodCompareCor[1:30,1:30])
```

As expected the methods are highly correlated.

### Plot the percentage of the variance explained:
```{r}
plot(1:length(funcPCA$sdev), cumsum(funcPCA$sdev^2/2346),
     ylab = '% Variance explained', 
     xlab = '# of principal of components',
     main = 'Method: PCA on Pearson correlations',
     ylim = c(0,1))
abline(v = mean(c(tail(which(cumsum(funcPCA$sdev^2/2346)<.8),1),
                  which(cumsum(funcPCA$sdev^2/2346)>.8)[1])))
plot(1:length(distPCA$sdev), cumsum(distPCA$sdev^2/2346),
     ylab = '% Variance explained', 
     xlab = '# of principal of components',
     main = 'Method: PCA on dissimilarity measure',
     ylim = c(0,1))
abline(v = mean(c(tail(which(cumsum(distPCA$sdev^2/2346)<.8),1),
                  which(cumsum(distPCA$sdev^2/2346)>.8)[1])))
plot(1:length(funcPCA$sdev), cumsum(funcPCA$sdev^2/2346),
     ylab = '% Variance explained', 
     xlab = '# of principal of components',
     main = 'Method: PCA on Pearson correlations',
     ylim = c(0,1))
abline(v = mean(c(tail(which(cumsum(funcPCA$sdev^2/2346)<.8),1),
                  which(cumsum(funcPCA$sdev^2/2346)>.8)[1])))
```

To explain 80% of the variance, both methods required around 33 dimensions (far fewer than 2346).

### Projection plots
```{r, echo=FALSE}
corPCAlbl = readRDS('../data/labels/funcConnlbls.rds')
distPCAlbl = readRDS('../data/labels/funcDistlbls.rds')
plot(scale(funcPCA$x[,1]),scale(funcPCA$x[,2]), 
     col = 1+(as.numeric(as.factor(corPCAlbl$GROUP))^2), xlab = 'dimension 1', ylab = 'dimension 2')
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
points(scale(distPCA$x[,1]),scale(distPCA$x[,2]), 
       col = 1+(as.numeric(as.factor(distPCAlbl$GROUP))^2),
     pch = 2, xlab = 'dimension 1', ylab = 'dimension 2')
legend('bottomleft', legend = c("Cor","Dist"), pch = 1:2)
title(main = "Cross-method similarity between Principal Components1 and 2")
```

This plot shows where the participants have ended up (their "PC scores") along dimensions 1 and 2. The circles represent those projections, one for each participant, as a result of conducting PCA on the pearson correlations, while the triangle-shaped points are the projections for the same individuals, but when the PCs are computed from the dissimilarity metric.

# Assess the variable loadings on each component
```{r}
dim1loadings = data.frame("PC1" = funcPCA$rotation[,1],
                          "var" = names(funcPCA$rotation[,1]),
                          stringsAsFactors = FALSE) %>%
  arrange(desc(abs(PC1)))
dim2loadings=data.frame("PC2" = funcPCA$rotation[,2],
                          "var" = names(funcPCA$rotation[,2]),
                          stringsAsFactors = FALSE) %>%
  arrange(desc(abs(PC2)))
head(dim1loadings[1:10,])
head(dim2loadings[1:10,])
```

```{r, eval=FALSE, echo=FALSE}
#python
#from numpy import genfromtxt
#load = genfromtxt('../results/PCA/cPCAloadings.csv', delimiter=',')
```

# Structural data PCA
```{r}
# read in the data and the list of variables to include
rawStructuralData = read.csv("../data/raw/structuralBrainData.csv") %>%
  mutate_at(vars("Measure.volume"), as.character)
newnm = names(rawStructuralData)
newnm[grep("Measure.volume", newnm)] = "SUBJECTID_long" 
names(rawStructuralData) <- newnm
varListStructural = read.csv("../data/structuralVarNames20191002_MVT.csv")
varsToKeep = as.character(varListStructural$var[!varListStructural$X...don.t.include == "X"])
# get labelled structural data and select only the correct variables
structLabels = readRDS("../data/structuralRaw+Labels+OverlappingVars.rds")[1:10]
# filter the raw data by choosing the wave1 data:
# also select the columns in varsToKeep
structureData = rawStructuralData[, varsToKeep]
structuralRawDataWithNames = merge(structLabels, rawStructuralData,
                                   by = "SUBJECTID_long",all.x = TRUE, all.y = FALSE)
saveRDS(structuralRawDataWithNames, "../data/processed/wave1structuralDataLABELLED.rds")
structuralRawData = structuralRawDataWithNames[,-1:-10]
```

# Run PCA
```{r}
brainStructurePCA = prcomp(structuralRawData, 
                           scale. = TRUE, center = TRUE, retx = TRUE)
saveRDS(brainStructurePCA, "../results/PCA/structPCAobj.rds")
```

#add the labels back in
```{r}
structPCAScores = data.frame(
  labels[keepRows, c("IDENT_SUBID","age","GROUP")],
  brainStructurePCA$x
)
saveRDS(structPCAScores, "../data/structPCAScoresLabelled.rds")
```

#Plots
### Variance-Explained
```{r}
plot(1:length(brainStructurePCA$sdev), cumsum(brainStructurePCA$sdev^2/24),
     ylab = '% Variance explained', 
     xlab = '# of principal of components',
     main = 'Method: PCA on Pearson correlations',
     ylim = c(0,1))
abline(v = mean(c(tail(which(cumsum(brainStructurePCA$sdev^2/24)<.8),1),
                  which(cumsum(brainStructurePCA$sdev^2/24)>.8)[1])))
```

### Projection plots
```{r}
par(mfrow = c(2,2))
plot(structPCAScores[,"PC1"], structPCAScores[,"PC2"],
     col = 1+(as.numeric(as.factor(structPCAScores$GROUP))^2), 
     xlab = 'dimension 1', ylab = 'dimension 2', pch = 16, cex = 1.5)
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
title("PIs and Comps projected onto PC 1 & 2")
plot(structPCAScores[,"PC1"], structPCAScores[,"PC3"],
     col = 1+(as.numeric(as.factor(structPCAScores$GROUP))^2), 
     xlab = 'dimension 1', ylab = 'dimension 3', pch = 16, cex = 1.5)
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
title("PIs and Comps projected onto PC 1 & 3")
plot(structPCAScores[,"PC2"], structPCAScores[,"PC3"],
     col = 1+(as.numeric(as.factor(structPCAScores$GROUP))^2), 
     xlab = 'dimension 2', ylab = 'dimension 3', pch = 16, cex = 1.5)
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
title("PIs and Comps projected onto PC 2 & 3")
plot(structPCAScores[,"PC1"], structPCAScores[,"PC4"],
     col = 1+(as.numeric(as.factor(structPCAScores$GROUP))^2), 
     xlab = 'dimension 1', ylab = 'dimension 4', pch = 16, cex = 1.5)
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
title("PIs and Comps projected onto PC 1 & 4")
```

