---
title: "sbDataWrangling"
author: "Ian Douglas"
date: "9/24/2019"
output:
  html_document:
    number_sections: yes
    toc: yes
    df_print: paged
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
    df_print: paged
---
```{r}
library(tidyverse)
```

# I. Load in the SB data.
#### The data contains 
* Structural MRI data for various brain regions
* Behavioral and demographic surveys
```{r}
load( # the .Rdata file is named "monster_SB"
  '../data/raw/3_ALL_long_cort_symptoms_brain_structure_function_epi_ages4-19_2019-05-15.Rdata'
)
```
#### II. Load in the time series folder names, in which one participant's TS exists
* Loading the folder names provides a list of TS data we have.
* A participants' TS's from different waves (1, 2, & 3) will be in different folders.
```{r}
subjects = list.files('../data/raw/raw_TS_by_subj')
head(subjects)
```
#### III. Using the above list and the master data ('monster_SB'):
* Create a data frame with grouping labels, basic demog., etc.
```{r}
lbls = data.frame(
  'IDENT_SUBID' = unique(substr(subjects, 1, 5)), stringsAsFactors = FALSE
) %>%
  # Record whether they have TS data at each wave
  group_by(IDENT_SUBID) %>%
  mutate(has_wave1_TS = 
           any(subjects %in% paste0(IDENT_SUBID,"_", IDENT_SUBID)),
         has_wave2_TS = 
           any(subjects %in% paste(rep(paste0(IDENT_SUBID,"_fu1"),2),collapse="_")),
         has_wave3_TS = 
           any(subjects %in% paste(rep(paste0(IDENT_SUBID,"_fu2"),2),collapse="_"))) %>%
  # Record which wave containing a TS is earliest
  mutate(wave_to_pull = min(which(c(has_wave1_TS,has_wave2_TS,has_wave3_TS)))) %>% 
  ungroup %>%
  #if a participant has only the third wave, delete them
  filter(wave_to_pull != 3) %>% 
  group_by(IDENT_SUBID) %>%
  #make a key to sort by later
  mutate(uniq_key = paste0(IDENT_SUBID,"_",wave_to_pull)) %>% ungroup

#Now merge in the ages
lblDataToPull = monster_SB %>% group_by(SUBJECTID_long) %>%
  mutate(uniq_key = 
           #this 'unique key' will distinguish the same subject at different waves
           paste0(sub("_.+$","",SUBJECTID_long),"_", index_wave)) %>%
  ungroup %>%
  select(uniq_key, age, GROUP) %>%
  merge(lbls, ., by = "uniq_key", all.x = TRUE)

#Create a variable that reformats the subjects IDs into the folder names containing their TS
lblDataToPull$folder_name = NA_character_
for (i in 1:nrow(lblDataToPull)) {
  if (lblDataToPull$wave_to_pull[i] == 2) {
    # these 'paste' lines just create the folder name, as named on the server
    fnm = paste(
      rep(
        paste0(
          lblDataToPull$IDENT_SUBID[i], "_fu", lblDataToPull$wave_to_pull[i] - 1
        ), times = 2
      ), collapse = "_"
    )
    lblDataToPull$folder_name[i] = fnm
  } else
    fnm = paste(
      rep(lblDataToPull$IDENT_SUBID[i],times=2), collapse = "_"
    )
  lblDataToPull$folder_name[i] = fnm
}

# Final notes:
# (a) Four ages are missing but these can be imputed later based on known data
# (b) some GROUP values are missing so input them now from the participant's other wave data:
for (i in filter(lblDataToPull,is.na(GROUP))$IDENT_SUBID) {
 tmp_dat = filter(monster_SB, IDENT_SUBID == i & !is.na(GROUP))
 lblDataToPull[lblDataToPull$IDENT_SUBID == i,"GROUP"] <- tmp_dat[,"GROUP"][1]
}
```
### Result:
```{r}
head(lblDataToPull)
saveRDS(lblDataToPull, '../data/labels/wave1Labels.rds')
```




