---
title: "structuralDataPCA"
author: "Ian Douglas"
date: "9/30/2019"
output:
  html_document:
    number_sections: yes
    toc: yes
    df_print: paged
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
    df_print: paged
---
```{r}
require(tidyverse)
```

# Read in the wave1 labels
```{r, eval=FALSE}
# (1) the labels for the participants, determining which wave to grab for each subject
lblDataToPull = readRDS("~/DANL/SB/wave1labels.rds")
# (2) the names of the structural variables we wish to retain
structVarNames = read.csv("~/DANL/SB/data/structuralVarNames20191002_MVT.csv",
                          stringsAsFactors = FALSE)
### and filter the ones we are going to keep.
strVarNamesToKeep = structVarNames %>%
  filter(X...don.t.include != "X")
# read in raw structural data
strRaw = read.csv(
  "~/DANL/SB/data/raw/structural/v6_asegstats_all_compiled_Marta_02-01-2018.csv",
  stringsAsFactors = FALSE) %>%
  # reformat subject id
  mutate_at("Measure.volume", ~sub("_FU","_fu",.))
# rename "Measure.volume" to match "SUBJECTID_long"
names(strRaw) = c("SUBJECTID_long", names(strRaw)[-1])

# Now retain the desired structural data
filteredStrRaw = strRaw[c("SUBJECTID_long",structVarstoKeep$var)] %>%
  # After discussion with Michelle, remove two more variables (bilateral ihVolumes)
  select(-rhCortexVol, -lhCortexVol)
```

### get the cbcl in the same manner as in the "rfModels.rmd"
### and loop in the sructural data
```{r, eval=FALSE}
# read in the data frame containing the cbcl
load( # the .Rdata file is named "monster_SB"
  '../data/raw/3_ALL_long_cort_symptoms_brain_structure_function_epi_ages4-19_2019-05-15.Rdata'
)
lblStrData = lblDataToPull[,c("IDENT_SUBID","SUBJECTID_long","age","GROUP","wave_to_pull")] %>%
  data.frame(., matrix(rep(NA,times=nrow(.)*(ncol(filteredStrRaw)-1)),
                       ncol = ncol(filteredStrRaw)-1)) %>%
  mutate(cbcl_totprob_t = NA)
names(lblStrData) = c(c("IDENT_SUBID","SUBJECTID_long","age","GROUP","wave_to_pull"),
                      names(filteredStrRaw)[-1], "cbcl_totprob_t")
for (i in 1:nrow(lblStrData)) {
  cbcl.dat <- monster_SB[grep(lblStrData$IDENT_SUBID[i],monster_SB$SUBJECTID_long),
                        c("CBCL_4_TOTPROB_T", "CBCL_4_TOTPROB_T_FU1")]
  if (all(is.na(cbcl.dat))) {
    lblStrData$cbcl_totprob_t[i] = NA
  } else
    if (lblStrData$wave_to_pull[i] == 1 & any(!is.na(cbcl.dat$CBCL_4_TOTPROB_T))) {
      lblStrData$cbcl_totprob_t[i] = na.omit(cbcl.dat$CBCL_4_TOTPROB_T)[1]
    } else if (nrow(cbcl.dat) > 1) {
        lblStrData$cbcl_totprob_t[i] = na.omit(cbcl.dat$CBCL_4_TOTPROB_T_FU1)[1]}
    if (lblStrData$wave_to_pull[i] == 2 & any(!is.na(cbcl.dat$CBCL_4_TOTPROB_T_FU1))) {
      lblStrData$cbcl_totprob_t[i] = na.omit(cbcl.dat$CBCL_4_TOTPROB_T_FU1)[1]
    } else if (nrow(cbcl.dat) > 1) {
        lblStrData$cbcl_totprob_t[i] = na.omit(cbcl.dat$CBCL_4_TOTPROB_T)[1]}
  if (lblStrData$SUBJECTID_long[i] %in% filteredStrRaw$SUBJECTID_long) {
    lblStrData[i, -c(1:5,ncol(lblStrData))] <-
      filteredStrRaw[filteredStrRaw$SUBJECTID_long == lblStrData$SUBJECTID_long[i], -1]
  }
}
```

# Note, filteredStrRaw retains whole brain vol (last variable), but all analyses should be run once with it, and once without.
```{r, eval=FALSE}
saveRDS(lblStrData, "~/DANL/SB/data/processed/structuralLabelled.rds")
saveRDS(select(lblStrData,-EstimatedTotalIntraCranialVol),
        "~/DANL/SB/data/processed/structLabelled_sansWholeBrainVol.rds")
```

# PCA on the structural data
```{r, eval=FALSE}
# select variables, then filter NA to retain maximum number of complete cases:
lblStrPCArawData = na.omit(lblStrData[c(1,6:27)]) # (including ident_subid)
lblStrPCArawData.noWBV = na.omit(lblStrData[c(1,6:26)]) # (no whole brain vol)
strPCA = prcomp(x = lblStrPCArawData[-1],
                scale. = TRUE, center = TRUE,
                retx = TRUE)
# and without whole brain volume
strPCA.noWBV = prcomp(x = lblStrPCArawData.noWBV[-1],
                scale. = TRUE, center = TRUE,
                retx = TRUE)
saveRDS(strPCA, "~/DANL/SB/results/PCA/structPCAobj.rds")
saveRDS(strPCA.noWBV, "~/DANL/SB/results/PCA/strPCAobj_noWBV.rds")
# now merge back in the other variables (which may have had NA)
saveRDS(left_join(
  cbind(lblStrPCArawData[1], strPCA$x),
  lblStrData[c("IDENT_SUBID","SUBJECTID_long","GROUP","age","cbcl_totprob_t")]),
        "~/DANL/SB/data/processed/strPCAscoresLabelled.rds")
saveRDS(left_join(
  cbind(lblStrPCArawData.noWBV[1], strPCA.noWBV$x),
  lblStrData[c("IDENT_SUBID","SUBJECTID_long","GROUP","age","cbcl_totprob_t")]),
        "~/DANL/SB/data/processed/strPCAscoresLabelled_noWBV.rds")
```

```{r, echo=FALSE}
strPCA = readRDS("~/DANL/SB/results/PCA/structPCAobj.rds")
strPCA.noWBV = readRDS("~/DANL/SB/results/PCA/strPCAobj_noWBV.rds")
```

# Evaluate fit
```{r}
par(mfrow = c(1,2))
plot(1:length(strPCA$sdev), cumsum(strPCA$sdev^2/ncol(strPCA$x)),
     ylab = 'Cumulative % Variance explained', 
     xlab = '# of principal of components',
     main = 'PCA on Structural Data',
     ylim = c(0,1))
abline(v = mean(c(tail(which(cumsum(strPCA$sdev^2/ncol(strPCA$x))<.8),1),
                  which(cumsum(strPCA$sdev^2/ncol(strPCA$x))>.8)[1])))
plot(1:length(strPCA.noWBV$sdev), cumsum(strPCA.noWBV$sdev^2/ncol(strPCA.noWBV$x)),
     ylab = 'Cumulative % Variance explained', 
     xlab = '# of principal of components',
     main = 'PCA on Structural Data \nwithout Cortical Volume',
     ylim = c(0,1))
abline(v = mean(c(tail(which(cumsum(strPCA.noWBV$sdev^2/ncol(strPCA.noWBV$x))<.8),1),
                  which(cumsum(strPCA.noWBV$sdev^2/ncol(strPCA.noWBV$x))>.8)[1])))
```

# Raw variable loading plots onto PCs
```{r, eval=FALSE, echo=FALSE}
loadingsList = lapply(
  1:7,
  function(x) {
    df = data.frame("loading" = strPCA$rotation[,x]) %>%
    mutate(
      "variable" = 
        factor(rownames(strPCA$rotation),
               levels =
                 rownames(strPCA$rotation)[order(abs(loading))]))
  df$SD = sapply(lblStrPCArawData[,as.character(df$variable)], sd)
  return(df)
  }
)
loadingPlots = lapply(
  1:6,
  function(x)
    ggplot(data = loadingsList[[x]]) +
    geom_bar(stat = "identity", 
             aes(x = variable, y = loading, fill = SD)) +
    coord_flip() +
    ggtitle(label = paste0("Loading onto PC", x))
    
)
# output plots
for (i in 1:6) {
  ggsave(paste0("~/DANL/SB/results/PCA/plots/STR_loadingsOntoPC",i,".pdf"), 
         plot = loadingPlots[[i]],
         height = 7, width = 7, units = "in", device = "pdf")
}
# for the version without whole-brain vol
loadingsList.noWBV = lapply(
  1:7,
  function(x) {
    df = data.frame("loading" = strPCA.noWBV$rotation[,x]) %>%
    mutate(
      "variable" = 
        factor(rownames(strPCA.noWBV$rotation),
               levels =
                 rownames(strPCA.noWBV$rotation)[order(abs(loading))]))
  df$SD = sapply(lblStrPCArawData.noWBV[,as.character(df$variable)], sd)
  return(df)
  }
)
loadingPlots.noWBV = lapply(
  1:6,
  function(x)
    ggplot(data = loadingsList.noWBV[[x]]) +
    geom_bar(stat = "identity", 
             aes(x = variable, y = loading, fill = SD)) +
    coord_flip() +
    ggtitle(label = paste0("Loading onto PC",x,"\nwithout Cranial Vol."))
    
)
# output plots
for (i in 1:6) {
  ggsave(paste0("~/DANL/SB/results/PCA/plots/STR_noWBV_loadingsOntoPC",i,".pdf"), 
         plot = loadingPlots.noWBV[[i]],
         height = 7, width = 7, units = "in", device = "pdf")
}
```

```{r}
loadingPlots.noWBV[[1]]; loadingPlots.noWBV[[2]]; loadingPlots.noWBV[[3]]; loadingPlots.noWBV[[4]]
```


# Projection plots
```{r, echo=FALSE, eval=FALSE, include=FALSE}
CoLoRs = 1+(as.numeric(as.factor(
  left_join(lblStrPCArawData["IDENT_SUBID"],lblStrData[c("IDENT_SUBID","GROUP")])$GROUP))^2)
jpeg("../results/PCA/plots/PC1+PC2structural.jpg",
     height = 6, width = 9, units = "in", res = 128)
plot(strPCA$x[,1],strPCA$x[,2], 
     col = CoLoRs, 
     xlab = 'dimension 1', ylab = 'dimension 2', pch = 16,
     main = "Data Projected onto PC1 and PC2")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
dev.off()
jpeg("../results/PCA/plots/PC1+PC3structural.jpg",
     height = 6, width = 9, units = "in", res = 128)
plot(strPCA$x[,1],strPCA$x[,3], 
     col = CoLoRs, 
     xlab = 'dimension 1', ylab = 'dimension 3', pch = 16,
     main = "Data Projected onto PC1 and PC3")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
dev.off()
jpeg("../results/PCA/plots/PC2+PC3structural.jpg",
     height = 6, width = 9, units = "in", res = 128)
plot(strPCA$x[,2],strPCA$x[,3], 
     col = CoLoRs, 
     xlab = 'dimension 2', ylab = 'dimension 3', pch = 16,
     main = "Data Projected onto PC2 and PC3")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
dev.off()
```

# Projection plots without whole-brain volume PCA
```{r, eval=FALSE, include=FALSE, echo=FALSE}
CoLoRs.noWBV = 1+(as.numeric(as.factor(
  left_join(lblStrPCArawData.noWBV["IDENT_SUBID"],
            lblStrData[c("IDENT_SUBID","GROUP")])$GROUP))^2)
jpeg("../results/PCA/plots/PC1+PC2structural_noWBV.jpg",
     height = 6, width = 9, units = "in", res = 128)
plot(strPCA.noWBV$x[,1],strPCA.noWBV$x[,2], 
     col = CoLoRs.noWBV, 
     xlab = 'dimension 1', ylab = 'dimension 2', pch = 16,
     main = "Data Projected onto PC1 and PC2\nwithout Brain Vol.")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
dev.off()
jpeg("../results/PCA/plots/PC1+PC3structural_noWBV.jpg",
     height = 6, width = 9, units = "in", res = 128)
plot(strPCA.noWBV$x[,1],strPCA.noWBV$x[,3], 
     col = CoLoRs.noWBV, 
     xlab = 'dimension 1', ylab = 'dimension 3', pch = 16,
     main = "Data Projected onto PC1 and PC3\nwithout Brain Vol.")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
dev.off()
jpeg("../results/PCA/plots/PC2+PC3structural_noWBV.jpg",
     height = 6, width = 9, units = "in", res = 128)
plot(strPCA.noWBV$x[,2],strPCA.noWBV$x[,3], 
     col = CoLoRs.noWBV, 
     xlab = 'dimension 2', ylab = 'dimension 3', pch = 16,
     main = "Data Projected onto PC2 and PC3\nwithout Brain Vol.")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
dev.off()
```

```{r}
plot(strPCA.noWBV$x[,1],strPCA.noWBV$x[,2], 
     col = CoLoRs.noWBV, 
     xlab = 'dimension 1', ylab = 'dimension 2', pch = 16,
     main = "Data Projected onto PC1 and PC2\nwithout Brain Vol.")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))

plot(strPCA.noWBV$x[,1],strPCA.noWBV$x[,3], 
     col = CoLoRs.noWBV, 
     xlab = 'dimension 1', ylab = 'dimension 3', pch = 16,
     main = "Data Projected onto PC1 and PC3\nwithout Brain Vol.")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))

plot(strPCA.noWBV$x[,2],strPCA.noWBV$x[,3], 
     col = CoLoRs.noWBV, 
     xlab = 'dimension 2', ylab = 'dimension 3', pch = 16,
     main = "Data Projected onto PC2 and PC3\nwithout Brain Vol.")
legend('topright', legend = c("PI","COMP"), fill = c(5,2))
```


